---
title: "Economic Analyses of Applicants"
author: "Joy Payton"
date: "3/15/2019"
output: 
  html_document: 
    toc: yes
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Load Packages

We'll need some packages to make this stuff work!

```{r}
library(dplyr)
library(stringr)
```


## Get Data

Let's bring in cat and dog application data.

```{r}
dog_applicants <- read.csv('../../Data/dog_apps.csv', stringsAsFactors = FALSE)
cat_applicants <- read.csv('../../Data/cat_apps.csv', stringsAsFactors = FALSE)
```

Also, what about the cards?

```{r}
dog_cards <- read.csv('../../Data/dog_cards.csv', stringsAsFactors = FALSE)
cat_cards <- read.csv('../../Data/cat_cards.csv', stringsAsFactors = FALSE)
```

And actions?

```{r}
dog_actions <- read.csv('../../Data/dog_actions.csv', stringsAsFactors = FALSE)
cat_actions <- read.csv('../../Data/cat_actions.csv', stringsAsFactors = FALSE)
```

As a reminder, in "Identify Census Tracts.Rmd" we created some data for the census tracts represented by our adopters, which contains selected economic characteristics.  Let's bring that data in!

```{r}
selected_economic_characteristics <- read.csv("CensusData/census_tracts_economics.csv", stringsAsFactors = FALSE)
```

Great, now we can do things like merge that selected data into our cat and dog applicants!

```{r}
cat_applicants <- merge(x = cat_applicants, 
                        y = selected_economic_characteristics,
                        by = "GEOID",
                        all.x = TRUE)
dog_applicants <- merge(x = dog_applicants, 
                        y = selected_economic_characteristics,
                        by = "GEOID",
                        all.x = TRUE)
```

## Analyzing Our Adopters' Financial Situation

Let's take a quick box plot of the median household income of our cat adopters (understood in aggregate):

```{r}
library(ggplot2)
ggplot(cat_applicants,aes(y=tract_median_income)) +
  geom_boxplot() 
```

And our dog adopters:

```{r}
ggplot(dog_applicants,aes(y=tract_median_income)) +
  geom_boxplot() 
```

In order to combine plots and compare cats vs dogs side-by-side, we should combine our data:

```{r}
all_applicants <- rbind(cat_applicants %>% mutate(pet = "cat"), 
                      dog_applicants %>% mutate(pet = "dog"))
```

Now we can do a side by side box plot:

```{r}
ggplot(all_applicants,aes(x=pet, y=tract_median_income)) +
  geom_boxplot() 
```

It seems like dog applicants come from slightly more affluent areas.  Does that bear up under statistical scrutiny?

```{r}
t.test(tract_median_income ~ pet, all_applicants)
t.test(tract_poverty_pct ~ pet, all_applicants)

```
**Yes, dog applicants are from wealthier neighborhoods, by a little (a few thousand dollars per year in income and handful of percent difference in the number of households under the poverty level).**

## Application Completion Status

How many applications were completed?  Does completion status have a link with median income?  With poverty?

First, let's combine cards with applications:

```{r}
cards <- rbind(cat_cards,dog_cards)
all_applicants <- merge(x = all_applicants,
                        y = cards, 
                        by.x ="outcome_trello_id",
                        by.y = "id")

```

Now we can plot application completion by various economic realities:

```{r}
ggplot(all_applicants,aes(x=dueComplete, y=tract_median_income)) +
  geom_boxplot() 

ggplot(all_applicants,aes(x=dueComplete, y=tract_poverty_pct)) +
  geom_boxplot() 
```

There does seem to be a small economic difference between completed applications -- applicants who completed come from slightly more affluent areas.  Is this difference statistically significant?

```{r}
t.test(tract_median_income ~ dueComplete, all_applicants)
t.test(tract_poverty_pct ~ dueComplete, all_applicants)
```

No, at an alpha of 0.05, this small difference is not statistically significant.

## Application Red Flags

How many applications got red flagged?  What is the economic situation there?

There are lots of labels, let's simplify, in order to make red flag, approvals, and denials easier to determine.

```{r}
all_applicants <- all_applicants %>% mutate(red_flag = str_detect(label_names, "red flag"),
                                              approved = str_detect(label_names, "approved"),
                                              denied = str_detect(label_names,"denied"))
```

What economic differences, if any, can be determined with regard to red flags?

```{r}
ggplot(all_applicants,aes(x=red_flag, y=tract_median_income)) +
  geom_boxplot() 
```

It certainly seems that applicants from poorer areas are more likely to receive red flags.  This may be completely desired, if the red flag is related to the financial ability to care for a pet.  But it might reflect bias.  How can we tell?

Let's start by looking at the budget set aside for the red flag holders.  We'll split out cats and dogs, since the two types of animals have different expected expenses.  First, let's convert currency to numbers:

```{r}
all_applicants$budget_monthly_num <- as.numeric(gsub('[$,]','',all_applicants$budget_monthly))
all_applicants$budget_emergency_num <- as.numeric(gsub('[$,]','',all_applicants$budget_emergency))
```

Let's see if our numbers make sense:

```{r}
summary(all_applicants$budget_monthly_num)
summary(all_applicants$budget_emergency_num)
```

Snap, we have some negative numbers and a couple of REALLY HIGH numbers.  Let's investigate.

```{r}
all_applicants %>% filter(budget_emergency_num < 0 |
                          budget_emergency_num > 10000 |
                          budget_monthly_num < 0 |
                          budget_monthly_num > 10000) %>%
                   select(outcome_trello_id, budget_monthly, budget_monthly_num, budget_emergency, budget_emergency_num)

```

Hmm.  I'm going to make a wild guess that negative numbers are a mistake and should be made positive.   I also suspect that things like \$0.50 are user errors and might be \$50.00, but I won't fix that.  Also, I wonder if some people put in their yearly salary or something like that instead of budget?

To try to resolve these, I'm going to do the following:

* Make negative numbers positive
* Exclude monthly budgets over $1,000 as unrealistic and probably an error
* Exclude emergency budgets over $50,000 as unrealistic and probably an error

```{r}
all_applicants$budget_monthly_num <- abs(as.numeric(gsub('[$,]','',all_applicants$budget_monthly)))
all_applicants$budget_emergency_num <- abs(as.numeric(gsub('[$,]','',all_applicants$budget_emergency)))
all_applicants$budget_emergency_num[which(all_applicants$budget_emergency_num > 50000)] <- NA
all_applicants$budget_monthly_num[which(all_applicants$budget_monthly_num > 1000)] <- NA
```

Let's check out the budget numbers...  First for monthly budgets:

```{r}
all_applicants %>%
  filter(!is.na(red_flag)) %>%
  ggplot(aes(x=pet, y=budget_monthly_num)) +
  geom_boxplot(aes(fill=red_flag)) 
```

Interestingly, there seems to be a strong monthly budget difference between cat red flag applicants and non red flag applicants.  But there's no appreciable difference in budget between dog owners based on red flag status.

Let's check that, statistically.  First, is the seemingly huge cat difference statistically significant?

```{r}
t.test(budget_monthly_num ~ red_flag, all_applicants %>% filter(!is.na(red_flag) & pet == "cat") )
```

Yes.  **There is a statistically significant difference in the monthly budgets proposed by cat applicants who received red flags and cat applicants who did not.**  This seems to indicate that unrealistic budgeting might come into play when red flags are imposed on cat applicants.

Dog applicants, on the other hand, seemed to have very similar monthly budgets, regardless of their red flag status.  Does that hold up?

```{r}
t.test(budget_monthly_num ~ red_flag, all_applicants %>% filter(!is.na(red_flag) & pet == "dog") )
```

There's no statistical significance to the very slight difference in budgets.

What about emergency budgets?

```{r}
all_applicants %>%
  filter(!is.na(red_flag)) %>%
  ggplot(aes(x=pet, y=budget_emergency_num)) +
  geom_boxplot(aes(fill=red_flag)) 
```

Snap, this is hard to check because of outliers.  Let's just set a ceiling at $10k for emergencies -- if a budget is more than that, we'll just set it to ten thousand.

```{r}
all_applicants$budget_emergency_num[which(all_applicants$budget_emergency_num > 10000)] <- 10000
all_applicants %>%
  filter(!is.na(red_flag)) %>%
  ggplot(aes(x=pet, y=budget_emergency_num)) +
  geom_boxplot(aes(fill=red_flag)) 
```

Looks like there's a difference here in both types of pets -- red flagged applicants budgeted less for emergencies.  Is this difference statistically significant?

```{r}
t.test(budget_emergency_num ~ red_flag, all_applicants %>% filter(!is.na(red_flag) & pet == "dog") )
t.test(budget_emergency_num ~ red_flag, all_applicants %>% filter(!is.na(red_flag) & pet == "cat") )
```

Yes, emergency budgets are lower for red-flagged applicants for both cats and dogs, at a statistically significant level.

## Budgets and Poverty
