---
title: "PAWS Initial Exploration"
author: "Kate Connolly"
output: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries & Files

```{r workspace, message=FALSE, warning=FALSE, results="hide"}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readxl)
library(readr)
library(zoo)
```

```{r readData, message=FALSE, warning=FALSE, results="hide"}
# data files
cat_apps <- read_csv("data/cat_apps.csv") %>%
  mutate(animal = "cat")
dog_apps <- read_csv("data/dog_apps.csv") %>%
  mutate(animal = "dog")

cat_actions <- read_csv("data/cat_actions.csv") %>%
  mutate(animal = "cat")
dog_actions <- read_csv("data/dog_actions.csv") %>%
  mutate(animal = "dog")

petpoint <- read_csv("data/petpoint.csv")

# combine
apps_combined <- rbind(cat_apps, dog_apps)
actions_combined <- rbind(cat_actions, dog_actions)
```
 

# Application Submission Dates

Made a plot to see the dates when applications were submitted, by `animal`. Not too much here. 

```{r echo=TRUE}

# first had to convert date_submitted from character to string (as.Date)
# then used as.yearmon to select the month and year only from the date

apps_combined %>%
  mutate(date_submitted_y_m = as.Date(as.yearmon(c(apps_combined$date_submitted), "%m/%d/%y"))) %>%
  ggplot(aes(x = date_submitted_y_m, fill = animal)) +
  geom_bar(position = "dodge")
```


# Monthly Budget Plots

```{r, echo=TRUE}

# convert monthly_budget from character to a numeric in order to do calculations
# removed parentheses, dollar signs, commas, negative sign
apps_combined$budget_monthly <- as.numeric(gsub('[()$,-]', '', apps_combined$budget_monthly))

# compute groupwise mean and median budget_monthly
apps_combined %>%
  group_by(animal) %>%
  summarize(mean(budget_monthly),
            median(budget_monthly))
```

Can see here that the median may be a more useful measure of center. The dog data is probably has more outliers that are skewing the mean, which can be further explored with a boxplot.

```{r echo=TRUE, message=FALSE, warning=FALSE}
# create box plot of budget_monthly by animal
apps_combined %>% 
  ggplot(aes(x = animal, y = budget_monthly)) +
  geom_boxplot()
```

Confirmed! Some allegedly very rich dog applicants are heavily influencing the mean. Removing some outliers and negative values gives a better picture.

```{r echo=TRUE, message=FALSE, warning=FALSE}
# create same box plot as above, with outliers and negative values removed
apps_combined %>%
  mutate(is_outlier = budget_monthly > 2000) %>%
  mutate(is_outlier_2 = budget_monthly < 0) %>%
  filter(!is_outlier) %>%
  filter(!is_outlier_2) %>%
  ggplot(aes(x = animal, y = budget_monthly)) +
  geom_boxplot()
  
```

There are still a few outliers, particularly in the cat data, but can see here that applicants for dogs tend to list a higher monthly budget. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# create hist of budget_monthly
apps_combined %>%
  ggplot(aes(x = budget_monthly)) +
  geom_histogram(bins = 15, color = "black", fill = "white") +
  xlim(c(0, 1000)) +
  facet_wrap(~ animal)
```

In general, both application sets had similar monthly budget distributions. Cat applications were a little more likely to list $100 or less compared to dog applications. But the median number, which was calculated above, shows a relatively similar distribution overall. 


Throwing a density plot in there too. Just like the histogram, can see the multiple modes. This could be due to a couple things — applicants may be unintentionally adding more zeros to their answers, or applicants are more likely to list numbers ending in zero. 

```{r echo=TRUE}
# same plot as above, just density plot
apps_combined %>%
  mutate(is_outlier = budget_monthly > 2000) %>%
  mutate(is_outlier_2 = budget_monthly < 0) %>%
  filter(!is_outlier) %>%
  filter(!is_outlier_2) %>%
  ggplot(aes(x = budget_monthly, fill = animal)) +
  geom_density(alpha = 0.3)
```
 

# Ideal Adoption Time & Adoption Dates

```{r echo=TRUE}
# bar plot for ideal_adoption_timeline by animal
ggplot(apps_combined, aes(x = ideal_adoption_timeline)) +
  geom_bar() +
  facet_wrap(~ animal) +
  coord_flip()
```

Cat applications listed "one-month-or-more" and "today" for ideal times. Dog applicants were similar, listing "few-weeks" and "today" most often. 

Also plotted the adoption outcome days by animal. 

```{r echo=TRUE}
# get all adoption outcome_date timestamps, for cats & dogs
# from petpoint, convert & select outcome_date & outcome_trello_id & animal_type and filter outcome_type by "Adoption"
outcome_dates <- petpoint %>%
  mutate(outcome_date = as.Date(outcome_date,"%m/%d/%y")) %>%
  select(outcome_type, outcome_date, outcome_trello_id, animal_type) %>%
  filter(outcome_type == "Adoption") %>%
  arrange(desc(outcome_trello_id))
  
# mutate removes time from outcome_date
outcome_dates %>%
#  mutate(outcome_date_day = date(outcome_dates$outcome_date)) %>%
  ggplot(aes(x = outcome_date)) +
  geom_bar() +
  facet_wrap(~ animal_type)
```

I didn't think the plot by day was very valuable, so then I wanted to see it grouped by month. Then it's clearer when more adoptions occurred — for cats, August & October through December; for dogs, a little more even distribution with some spikes in August & December.

The dates labels aren't in the right order, but it gets the job done for now.  

```{r echo=TRUE}
# mutate sets the outcome_date to just year and month, basically grouping counts by month
outcome_dates %>%
  mutate(outcome_month_yr = format(as.Date(outcome_dates$outcome_date), "%Y-%m")) %>%  
  ggplot(aes(x = outcome_month_yr)) +
  geom_bar() +
  facet_wrap(~ animal_type) +   # show one chart for dogs and one for cats
  coord_flip()   # flip the x and y axis
  
```
 

# Application to Adoption Duration

Wanted to try to take a look at the time between an application being submitted (in `cards`) and an adoption being finalized (in `petpoint`). Not sure if some of these methods are the correct/most efficient way to do this, but gave it a shot. 

```{r echo=TRUE}
## The goal is to find all of the starting times & all the ending times
## Start times can be found in the actions data, end times in PetPoint
## First want to organize only the data that we need before joining the two data sets


# get all the times when a card was created for cats
# from cat_actions, select date & id and filter by "createCard" type
create_card <- actions_combined %>%
  select(type, date, data.card.id) %>%
  filter(type == "createCard")

# check to see that there are no duplicate ids; run and consult console
nrow(distinct(create_card, data.card.id))
```

```{r echo=TRUE}
# get all adoption outcome_date timestamps, for cats & dogs
# from petpoint, select outcome_date & outcome_trello_id & animal_type and filter outcome_type by "Adoption"
outcome_dates <- petpoint %>%
  select(outcome_type, outcome_date, outcome_trello_id, animal_type) %>%
  filter(outcome_type == "Adoption") %>%
  arrange(desc(outcome_trello_id))
  
# check to see that there are no duplicate ids; run and consult console
nrow(distinct(outcome_dates, outcome_trello_id))
```

```{r echo=TRUE}
# saw that there were indeed duplicates (petpoint is the culprit)
# de-dupe
outcome_dates <- outcome_dates[!duplicated(outcome_dates$outcome_trello_id),]

# join the two datasets
start_and_end_dates <- merge(create_card, outcome_dates, by.x="data.card.id", by.y="outcome_trello_id")

# count the number of dogs & cats in the outcome_dates
count(outcome_dates, vars = animal_type)  
```
 
Next up is getting the dates and times in similar formats.

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
# make the relevant date columns date format
start_and_end_dates <-
  start_and_end_dates %>%
  mutate(outcome_date = as.Date(outcome_date,"%m/%d/%y" )) %>%
  mutate(date = as.Date(date, "%m/%d/%y"))

#start_and_end_dates <-
#  start_and_end_dates %>%
#  mutate(create_date_est = date(start_and_end_dates$date)) %>%    # remove hour from date EST time
#  mutate(final_date = date(start_and_end_dates$outcome_date))    # remove hour from outcome_date
  
# add column for difference between create and outcome dates (duration) 
start_and_end_dates <-
  start_and_end_dates %>%
  mutate(duration = start_and_end_dates$outcome_date - start_and_end_dates$date) 

# convert negative values in duration column to 0
start_and_end_dates <-
  start_and_end_dates %>%
  mutate(duration2 = ifelse(start_and_end_dates$duration < 0, 0, start_and_end_dates$duration))   
```

Historgram that shows the distribution of duration. Initially used `facet_wrap` to compare cat and dog, but I think it's easier to just look at them as separate plots. 

```{r echo=TRUE}
# initial histogram plot
gp1 <- start_and_end_dates %>%
      filter(animal_type == "Cat") %>%
      ggplot(aes(x = duration2)) +
      theme_bw() +
      geom_histogram(bins = 37, color = "black", fill = "lightgray") +
      ggtitle("Distribution of Cat App Duration (in days)") +
      theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
      xlab("Days Until Adoption") +
      ylab("Count of Apps") +
      geom_vline(aes(xintercept = mean(duration2), 
                     color = (blue = "Mean")),
                linetype = "dashed",
                size = 1,
                show.legend = TRUE) +
      geom_vline(aes(xintercept = median(duration2), 
                     color = (red = "Median")),
                 linetype = "dashed",
                 size = 1,
                 show.legend = TRUE)

# customize legend
gp1 +
  scale_color_manual(name = '', 
                     labels = c(Mean = "Mean", Median = "Median"),
                     values = c(Mean = 'blue', Median = 'red')) 
```


```{r echo=TRUE}
# initial histogram plot
gp1 <- start_and_end_dates %>%
      filter(animal_type == "Dog") %>%
      ggplot(aes(x = duration2)) +
      theme_bw() +
      geom_histogram(bins = 37, color = "black", fill = "lightgray") +
      ggtitle("Distribution of Dog App Duration (in days)") +
      theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
      xlab("Days Until Adoption") +
      ylab("Count of Apps") +
      geom_vline(aes(xintercept = mean(duration2), 
                     color = (blue = "Mean")),
                linetype = "dashed",
                size = 1,
                show.legend = TRUE) +
      geom_vline(aes(xintercept = median(duration2), 
                     color = (red = "Median")),
                 linetype = "dashed",
                 size = 1,
                 show.legend = TRUE)

# customize legend
gp1 +
  scale_color_manual(name = '', 
                     labels = c(Mean = "Mean", Median = "Median"),
                     values = c(Mean = 'blue', Median = 'red')) 
```

Perhaps not immediately obvious in the visualizaions, but doing some quick calculations shows that the duration for cats and dogs is pretty similar. 

```{r echo=TRUE}
start_and_end_dates %>%
  group_by(animal_type) %>%
  summarize(mean(duration2),
            median(duration2))
```

But there are a fair amoung of outliers. Those are what we'll want to examine to understand what makes an application quick or slow. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# create box plot of budget_monthly by animal
start_and_end_dates %>% 
  ggplot(aes(x = animal_type, y = duration2)) +
  geom_boxplot()
```
