---
title: "MergedDataAnalysis-Adoption"
author: "Ramaa Nathan"
date: "3/21/2019"
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
#  html_document
knitr::opts_chunk$set(echo = TRUE)
#options(knitr.table.format = "html") 
library(tidyverse)
library(lubridate)
library(skimr)
library(gtools)
library(knitr)  #for outputting tables as kables
library(caret)
```

```{r load}
master <- readRDS("../masterapps_20190324.rds")
dim(master)
#colnames(master)
#outcome_type contains the final status
#are there any NA?
anyNA(master$outcome_type)
#What are the unique labels?
unique(master$outcome_type)
# All are only adoption

#Find column names without ind
#master %>% select(-ends_with("_ind")) %>% colnames()
```

#### Check the labels of the entries that correspond to adoption for any irregular patterns
```{r }
#master %>% select(-contains("_ind")) %>% str()
#master %>% select(contains("_ind")) %>% str()

unique(master$last_label)
#master %>% select(trello_id, outcome_type,contains("label"))

#check if any of the labels do not contain adopted
master %>% 
  filter(outcome_type == "Adoption") %>% 
  filter(!str_detect(label_names,"adopted")) %>% 
  select(trello_id, outcome_type,label_names) 
## 28 entries found!

#check if any of the labels do not contain "adopt" in any form
master %>% 
  filter(outcome_type=="Adoption") %>% 
  filter(!str_detect(label_names,"adopt")) %>%
  select(trello_id, outcome_type,label_names) 
# there are four entries with no labels associated with label

#check if there are no labels at all
master  %>% 
  filter(outcome_type=="Adoption") %>% 
  filter(num_labels == 0) %>% 
  select(trello_id,animal_type,label_names)
# there are six total entries and fice unique entries with no labels

#master  %>% filter(outcome_type=="Adoption") %>% filter(num_labels == 0) %>%  select(trello_id,label_names,date_submitted,outcome_date,starts_with("checklist")) 
#master  %>% filter(outcome_type=="Adoption") %>% select(trello_id,starts_with("checklist")) 


```
#### Verify the animal_type column - a problem was detected here
```{r}
#check why we have animal.x and animal.y
master %>% 
  select(starts_with("animal_type")) %>% 
  table()
```

### Plots for adopted
```{r}
master %>% 
  filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=animal_type,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Species") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

master %>% 
  filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=specific_animal,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Specific Animal") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=new_age_group,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Age Group of Animal") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=budget_monthly_ranges,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Monthly budgets") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=budget_emergency_ranges,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Emergency budgets") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Home alone avg
master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=home_alone_avg,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Home Alone Average") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Home alone max
master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=home_alone_max,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Home Alone Maximum") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Adults at home
master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=adults_in_home,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Number of Adults in Home") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### CHildren at home
master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=children_in_home,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Number of Children in Home") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Experience
master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=experience,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Experience with Pets") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Home Pet Policy
master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=home_pet_policy,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Home Pet Policy") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Pet Return
master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=return_pet,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Pet Return") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Allergies
master %>% filter(outcome_type=="Adoption") %>%
  ggplot(mapping=aes(x=allergies,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Allergies") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

```

#### Apply Random Forests to check for important varaibles
```{r}
master_mod <- master %>%
  mutate(adoption = ifelse((!is.na(outcome_type) & outcome_type=="Adoption"),1,0)) %>%
  #mutate(adoption = ifelse((!is.na(outcome_type) & outcome_type==1),1,0)) %>%
  select(-c(trello_id,reason_for_adoption, all_household_agree, allergies, home_owner, all_household_agree, allergies,
            experience, budget_monthly, budget_emergency,pet_kept,exercise,needs,return_pet),
         -starts_with("budget."),
         -starts_with("home.alone"),
         -starts_with("checklist"),  #the type of check may not affect final adoption
         -starts_with("label"),  #labels are not reliable
         -starts_with("new.age"))

## find columns with NA
master_mod %>% skim() %>% 
  filter(stat == "missing") %>% filter(value > 0) %>% 
  kable(caption="Missing Entries in master_mod")
```

#### Impute the values
All _ind columns should have either a 1 or 0. There is some bug in the convert_to_ind funtion what was used during merge that is creating NA in the _ind columns. Set them to 0.

```{r}

#replace all _ind NAs with 0

master_mod <- master_mod %>%
  mutate_at(vars(ends_with("_ind")),
            funs(if_else(is.na(.), 0, .)))

```


#### Apply random forests
```{r}
#removing all NAs results in dataframe with 0 rows
dim(na.omit(master_mod))

master_pre_process <- preProcess(master_mod, "knnImpute")

master_pre_processpred <- predict(master_pre_process, master_mod)

inTrain <- createDataPartition(y = master_mod$adoption, p=0.7, list=FALSE)
training <- master_mod[inTrain, ]
testing <- master_mod[-inTrain,]
dim(training)
dim(testing)
table(training$adoption)
table(testing$adoption)
rfFit <- train(adoption ~ .,
               data = training,
               method = "rf",
               prox = TRUE,
               importance = TRUE)
#rffit
#importance(rffit)
```

