---
title: "Animal Intake Characteristics"
author: "Javier Jasso"
output:
  pdf_document: default
  html_document: default
code_folding: hide
---

```{r first chunk, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
```

##Research Question:
How did intake type (Surrender, Return, Stray, and Transfer in) affect application wait time (i.e., intake to release date)?

##Dataset 
```{r}
#load packages
library(tidyverse)
library(psych)

#import data
master_animal <- read.csv("~/Documents/GitHub/2019_datathon/Data/processed_data/master_animal.csv")

```

```{r data_manipulation}
#use dplyr to get subset of data
library(tidyverse)

master_animal_intake <- master_animal %>% 
    dplyr::select(intake_type,
           intake_subtype,
           wait_days, 
           animal_type) %>% 
    filter(!is.na(intake_type)) %>% 
    mutate(intake_type = factor(intake_type), 
           animal_type = factor(animal_type), 
           wait = as.numeric(wait_days))

table(master_animal_intake$intake_type)
table(master_animal_intake$intake_subtype)
```

```{r subtype}
subtype_filter <- master_animal_intake %>% 
    filter(intake_subtype == 'Returned Adoption' |
               intake_subtype == 'Public Drop Off')
```

## Descriptive Data
NB: one case removed for analysis due to missing intake type
NB: More NAs in the animal_type variable

We can summarize the data above with a table:
```{r}
master_animal_intake %>%
    group_by(intake_type) %>% 
    summarise(Wait = mean(wait_days),
              SD = sd(wait_days),
              n = n())
```
Now the same table, by animal type (cat vs. dog)
```{r summary_animal.type}
cats <- master_animal_intake %>%
    group_by(intake_type) %>% 
    filter(animal_type=='cat') %>% 
    summarise(time=mean(wait_days), SD=sd(wait_days), n=n())
cats
dogs <- master_animal_intake %>%
    group_by(intake_type) %>% 
    filter(animal_type == 'dog') %>% 
    summarise(time=mean(wait_days), SD=sd(wait_days), n=n())
dogs
```

And a table of Intake Subtype variable:  
```{r}
table(master_animal_intake$intake_subtype)

```
For this analysis, we retained 2 of the variables: Public Dropoff (n = 433) and Returned Adoption (n = 213).

##  Plotting
Frequency of "Intake Type" variable:
```{r}
#cleanup code
ggplot.cleanup = theme(panel.grid.major = element_blank(),
                       panel.grid.minor = element_blank(),
                       panel.background = element_blank(),
                       axis.line = element_line(color = 'black'))

intake_type_hist <- ggplot(master_animal_intake, aes(intake_type))
intake_type_hist +
    geom_bar(color="turquoise", fill="turquoise")+
    xlab('Intake Type') +
    ylab('Frequency')+
    ggplot.cleanup +
    theme(axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20)) +
    coord_flip()

```

Frequency of "Intake Type" variable:
```{r}
#plotting fequency of intake_subtype var:
intake_subtype_all_hist <- ggplot(master_animal_intake, aes(intake_subtype))
intake_subtype_all_hist +
    geom_bar(color="gold", fill="gold")+
    xlab('Intake Subtype') +
    ylab('Frequency')+
    ggplot.cleanup +
    theme(axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20)) +
    coord_flip()
```


```{r}

intake_subtype_hist <- ggplot(subtype_filter, aes(intake_subtype))
intake_subtype_hist +
    geom_bar(color="gold", fill="gold")+
    xlab('Intake Subtype') +
    ylab('Frequency')+
    ggplot.cleanup +
    theme(axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20)) +
    coord_flip()

```

#Analysis
Now let's run multiple regression models.  
Here, we have a model with Intake Type predicting Wait:
```{r eval = FALSE}
intake_type_lm <- lm(wait_days ~ intake_type, master_animal_intake)
summary(intake_type_lm)
```

A no-intercept version of the same variables:
```{r}
intake_type_lm_no.int <- lm(wait_days ~ intake_type - 1, master_animal_intake)
summary(intake_type_lm_no.int)

ggplot(master_animal_intake, aes(x = intake_type, y = wait_days)) +
    geom_point() +
    stat_smooth(method = "lm")
```

And, for fun, let's run an ANOVA using aov()
```{r}
intake_aov <- aov(wait_days ~ intake_type, master_animal_intake) 
summary(intake_aov)
```

Also, we can test for a possible Intake Type x Animal Type interaction:
```{r interaction}
intake_type_int <- lm(wait_days ~ intake_type*animal_type, master_animal_intake)
summary(intake_type_int)
```

Regression model with Public Dropoff and Returned Adoption:
```{r lm_subtype}
#lm subtype:
intake_subtype_lm <- lm(wait_days ~ intake_subtype, subtype_filter)
summary(intake_subtype_lm)
intake_subtype_lm$coefficients
##lm subtype plot:
ggplot(subtype_filter, aes(x = intake_subtype, y = wait_days)) +
    geom_point() +
    stat_smooth(method = "lm")
```

#Results: 
##Intake Type
There are significant differences in time, F(3, 2781) = 32.67, p < .001, such that "stray"" and "transfer" had significantly **longer** wait times than "surrender," while there were no significantly different wait times between "surrender" and "return."  

##Intake Subtype
Mean wait days: “public drop off” = 55; “returned adopt” = 32.  
There are significantly different wait, with “public drop off” > “returned adopt.”


