---
title: "PAWS - R Ladies Community Data Project"
author: "Ramaa Nathan"
date: "2/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#options(knitr.table.format = "html") 
library(tidyverse)
library(lubridate)
library(skimr)
library(gtools)
library(knitr)  #for outputting tables as kables
#library(kableExtra)  # for formatting kable
```

#### petpoint.csv
Summary of Transformations:
1. DOB: Convert from character to Date Type using : lubridate::mdy(dob) 

2. intake_date, release_date, outcome_date: convert to POSIX using mdy_hm(<date>)

3. Convert all other character types to factor types

4. Create time intervals and and time period caterogies to track time from intake to outcome.

Summary of data: using skimr::skim(petpoint)

Questions:
1. NAME: Regional name: Data is a number string. Are these IDS?

2. There are 12 entries with intake_age = -1 months and 312 entries with intake_age=0 months. What is the difference between them?

3. ZipCode: there are 41 entries with invalid zipcodes (length < 5)

4. There are 308 trello_ids where there is more than one duplicate of outcome_trello_id. On checking, we find 242 entries with 2 duplicates, 10 ids with three dupliates, 2 ids with 4 duplciates each, and 1 id with 5 duplicates


```{r petpoint,echo = FALSE}
petpoint = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/petpoint.csv")
paste("First read in:")
dim(petpoint)

# Are there any duplicates in the data?
which(duplicated(petpoint[,-1]))
#remove the duplicates
petpoint <- petpoint[-which(duplicated(petpoint[,-1])),]
paste("after removing duplicates")
dim(petpoint)

 # Are there any duplicates of outcome_trello_id where outcome_trello_id is  id for matching with cards and actions datasets
 numNonUniquesIDs = petpoint %>% filter(!is.na(outcome_trello_id)) %>% select(outcome_trello_id) %>% group_by(outcome_trello_id) %>% summarise(dupcount=n())
table(numNonUniquesIDs$dupcount)
#find the one with 5 duplicate outcome_trello_id
fivecount=numNonUniquesIDs %>% filter(dupcount == 5)
#get more details about animal details - outcome type and outcome date for this duplicat trello_id
petpoint %>% semi_join(fivecount) %>% select(outcome_trello_id,outcome_type, outcome_date) %>% arrange(outcome_date)

#find the one with 2 duplicate outcome_trello_id
twocount=numNonUniquesIDs %>% filter(dupcount == 2)
#get more details about animal details - outcome type and outcome date for this duplicat trello_id
petpoint %>% semi_join(twocount) %>% select(outcome_trello_id,species,gender,primary_breed,outcome_type, outcome_date) %>% arrange(outcome_trello_id, outcome_date)
```

#### Data Transformations
```{r,echo = FALSE}
#first convert the date columns and then convert the rest of the character columns to factors
petpoint <- petpoint %>% 
  mutate(dob=mdy(dob),
         intake_date=mdy_hm(intake_date,tz="America/New_York"),
         release_date=mdy_hm(release_date,tz="America/New_York"),
         outcome_date=mdy_hm(outcome_date,tz="America/New_York")
  )

petpoint <- petpoint %>% mutate_if(sapply(petpoint,is.character),as.factor)
str(petpoint)
skim(petpoint)
```

#### Missing data and data anomalies
```{r,echo = FALSE}
# find number of missing data in each column
print("Missing Entries")
petpoint %>% skim() %>% 
  filter(stat == "missing") %>% filter(value > 0) %>% 
  kable(caption="Missing Entries in Petpoint.csv")

#Find the entries that have incorrect format of zipcode
paste("Number of invalid zipcode entries:",petpoint %>%  filter(nchar(outcome_ZIP) < 5) %>% length())
# what is the distribution of invalid zipcodes
petpoint %>%  filter(nchar(outcome_ZIP) < 5) %>% select(outcome_ZIP) %>% table
#petpoint %>%  filter(nchar(outcome_ZIP) < 5) %>% select(outcome_ZIP) %>%
#  group_by(outcome_ZIP) %>% summarise(count=n()) %>%
  #kable(caption="Frequency of invalid zipcodes") %>%
  #kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

#are the intake age ranges valid?
table(petpoint$age_intake)
print("There are 12 entries with age = -1 months and 312 entries with age=0 months")

## This is to be sure that I did not introduce any duplicates
paste("After all processing:")
dim(petpoint)
```

#### Data Visualizations
```{r,echo = FALSE}
# reorder the age_group levels
petpoint <- petpoint %>% 
  mutate(age_group = factor(age_group,levels(age_group)[c(10,5,9,3,7,1,4,6,8,2)],ordered=TRUE))
petpoint %>% ggplot(mapping=aes(x=age_group,fill=age_group)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Age Group Distribution") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

#Distribution of species
 petpoint%>% ggplot(mapping=aes(x=species,fill=species)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Distribution of Species") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
#plot the intake asilomar
 petpoint%>% ggplot(mapping=aes(x=intake_asilomar,fill=intake_asilomar)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(title = "Intake Asilomar")  +
    geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
#plot the intake type
 petpoint%>% ggplot(mapping=aes(x=intake_type,fill=intake_type)) + 
    geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Intake Type")  +
    geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
 #plot the intake sub-type
 petpoint%>% ggplot(mapping=aes(x=intake_subtype,fill=intake_subtype)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   labs(title = "Intake Sub Type") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
 #plot the intake reason
 petpoint%>% ggplot(mapping=aes(x=intake_reason,fill=intake_reason)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(title = "Intake Reason") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 table(petpoint$intake_reason)
 
 #plot the intake sitename
 petpoint%>% ggplot(mapping=aes(x=intake_sitename,fill=intake_sitename)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Intake Sitename") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
  #plot the outcome sitename
 petpoint%>% ggplot(mapping=aes(x=outcome_sitename,fill=outcome_sitename)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Outcome Sitename") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

 #plot the outcome type
 petpoint%>% ggplot(mapping=aes(x=outcome_type,fill=outcome_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Outcome Type") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
  #plot the outcome sub-type
 petpoint%>% ggplot(mapping=aes(x=outcome_subtype,fill=outcome_subtype)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Outcome Sub Type") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
kable(sort(table(petpoint$outcome_subtype),decreasing=TRUE))
 
#How many were altered?
 petpoint%>% ggplot(mapping=aes(x=altered,fill=altered)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Sprayed/Neutered") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
 # How many days between intake and outcome 
petpoint <- petpoint %>% mutate(process_time = interval(intake_date,outcome_date) / ddays(1))
#plot the distribution
petpoint %>% ggplot(mapping=aes(x=process_time)) +
  geom_histogram(breaks=seq(20,300,by=20),col="red",fill="green",alpha=0.2) +
  labs(title="Histogram of Processing Time in Days (Intake -> Outcome)")

#create time period categories
petpoint$process_time_periods <- cut(petpoint$process_time,
      breaks=c(-Inf,1,3,5,10,30,90,180,Inf),
      labels=c("< 1day","2-3 days","4-5 days","6-10 days", "11-30 days", "31-90 days", "91-180days", ">180 days"))
#plot the categories
 ggplot(petpoint,mapping=aes(x=process_time_periods,fill=process_time_periods)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Process Time Periods") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

 
```



