---
title: "PAWS - R Ladies Community Data Project"
author: "Ramaa Nathan"
date: "2/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#options(knitr.table.format = "html") 
library(tidyverse)
library(lubridate)
library(skimr)
library(gtools)
library(knitr)  #for outputting tables as kables

```

#### petpoint.csv
Summary of Transformations:
1. DOB: Convert from character to Date Type using : lubridate::mdy(dob) 

2. intake_date, release_date, outcome_date: convert to POSIX using mdy_hm(<date>)

3. Convert all other character types to factor types

4. Create time intervals and and time period caterogies to track time from intake to outcome.

Summary of data: using skimr::skim(petpoint)

Questions:
1. NAME: Regional name: Data is a number string. Are these IDS?

2. There are 12 entries with intake_age = -1 months and 312 entries with intake_age=0 months. What is the difference between them?

3. ZipCode: there are 41 entries with invalid zipcodes (length < 5)

4. There are 308 trello_ids where there is more than one duplicate of outcome_trello_id. On checking, we find 242 entries with 2 duplicates, 10 ids with three dupliates, 2 ids with 4 duplciates each, and 1 id with 5 duplicates

#### Process the data file
```{r petpoint,echo = FALSE}
petpoint = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/petpoint.csv")
paste("First read in:")
dim(petpoint)

# Are there any duplicates in the data?
sum(duplicated(petpoint[,-1]))
#remove the duplicates
#petpoint <- petpoint[-which(duplicated(petpoint[,-1])),]
petpoint <- petpoint[,-1] %>% distinct()
paste("after removing duplicates")
dim(petpoint)

 # Are there any duplicates of outcome_trello_id where outcome_trello_id is  id for matching with cards and actions datasets
 numNonUniquesIDs = petpoint %>% 
   filter(!is.na(outcome_trello_id)) %>% 
   select(outcome_trello_id) %>% 
   group_by(outcome_trello_id) %>% 
   summarise(dupcount=n())
paste("Frequency of non unique trello IDs")
table(numNonUniquesIDs$dupcount)


```

#### Data Transformations
```{r,echo = FALSE}
#first convert the date columns and then convert the rest of the character columns to factors
petpoint <- petpoint %>% 
  mutate(dob=mdy(dob),
         intake_date=mdy_hm(intake_date,tz="America/New_York"),
         release_date=mdy_hm(release_date,tz="America/New_York"),
         outcome_date=mdy_hm(outcome_date,tz="America/New_York")
  )

#Find the entries that have incorrect format of zipcode
paste("Initial Number of invalid zipcode entries:",petpoint %>%  filter(nchar(outcome_ZIP) < 5) %>% select(outcome_ZIP) %>% dim())
petpoint %>%  filter(str_length(outcome_ZIP) < 5) %>% select(outcome_ZIP)

# convert zipcode from integer to character format and add a zero at the beginning when length is less than 5
petpoint <- petpoint %>% 
  mutate(outcome_ZIP=as.character(outcome_ZIP)) %>% 
  mutate(outcome_ZIP=ifelse(str_length(outcome_ZIP)<5,str_c("0",outcome_ZIP),outcome_ZIP))
  
#Recheck if there are still any  entries that have incorrect format of zipcode
paste("After correction: Number of invalid zipcode entries:",petpoint %>%  filter(str_length(outcome_ZIP) < 5) %>% select(outcome_ZIP) %>% dim())
petpoint %>%  filter(str_length(outcome_ZIP) < 5) %>% select(outcome_ZIP)

#convert any remaining columns from character type to factor type
petpoint <- petpoint %>% mutate_if(sapply(petpoint,is.character),as.factor)

str(petpoint)
skim(petpoint)
```

####  Checks for Missing data and data anomalies
```{r,echo = FALSE}
# find number of missing data in each column
print("Missing Entries")
petpoint %>% skim() %>% 
  filter(stat == "missing") %>% filter(value > 0) %>% 
  kable(caption="Missing Entries in Petpoint.csv")

#check primary breed names
paste( "Primary Breeds")
unique(petpoint$primary_breed)
paste("Secondary breeds:")
unique(petpoint$secondary_breed)
paste("Markings:")
unique(petpoint$markings)
paste("Species:")
unique(petpoint$species)

#are the intake age ranges valid?
table(petpoint$age_intake)
print("There are 12 entries with age = -1 months and 312 entries with age=0 months")


#find the one with 5 duplicate outcome_trello_id
fivecount=numNonUniquesIDs %>% 
  filter(dupcount == 5)

#get more details about animal details - outcome type and outcome date for this duplicat trello_id
petpoint %>% semi_join(fivecount) %>% 
  select(outcome_trello_id,outcome_type, outcome_date) %>% 
  arrange(outcome_date)

#find the one with 2 duplicate outcome_trello_id
twocount=numNonUniquesIDs %>% 
  filter(dupcount == 2)

#get more details about animal details - outcome type and outcome date for this duplicat trello_id
petpoint %>% semi_join(twocount) %>% 
  select(outcome_trello_id,species,gender,primary_breed,outcome_type, outcome_date) %>% 
  arrange(outcome_trello_id, outcome_date)


## This is to be sure that I did not introduce any duplicates
paste("After all processing:")
dim(petpoint)
```
```{r}
##Compare dob, intake_date, age_intake, age_group 
#petpoint %>% sample_n(10) %>% select(dob, intake_date, age_intake,age_group) 


petpoint <- petpoint %>%  select(-age_group) %>%
  mutate(new_age_group = factor(case_when(age_intake<=1 ~ "<4 weeks", 
                                   age_intake <= 3 ~ "4-12 weeks", 
                                   age_intake <= 6 ~ "12weeks-6months",
                                   age_intake <= 12 ~ "6months-1year",
                                   age_intake <= 24 ~ "1-2years",
                                   age_intake <= 48 ~ "2-4years",
                                   age_intake <= 72 ~ "4-6years",
                                   age_intake <= 120 ~ "6-10years",
                                   is.na(age_intake) ~ "NA",
                                   TRUE ~ "older than 10years"),
                                levels=c("<4 weeks","4-12 weeks","12weeks-6months",
                                         "6months-1year","1-2years","2-4years",
                                         "4-6years","6-10years","older than 10years","NA"),
                                ordered=TRUE))
#Verify that the new_agee group is correct - Compare dob, intake_date, age_intake, new_age_group 
petpoint %>% sample_n(10) %>% select(dob, intake_date, age_intake,new_age_group) 
```
Observation: When we compare the DOB, age_intake, intake_date and age_group, the age-group column does not match the intake_age. So, dropped the original age_group column and added a new_age_group column

#### Data Visualizations
```{r,echo = FALSE}
# reorder the age_group levels
#petpoint <- petpoint %>% 
   #mutate(age_group = factor(age_group,levels(age_group)[c(10,5,9,3,7,1,4,6,8,2)],ordered=TRUE))
petpoint %>% ggplot(mapping=aes(x=new_age_group,fill=new_age_group)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Age Group Distribution") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

#Distribution of species
 petpoint%>% ggplot(mapping=aes(x=species,fill=species)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Distribution of Species") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

 #Distribution of agency name
 petpoint%>% ggplot(mapping=aes(x=agency_name,fill=agency_name)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Distribution of Agency Name") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
#plot the intake asilomar
 petpoint%>% ggplot(mapping=aes(x=intake_asilomar,fill=intake_asilomar)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(title = "Intake Asilomar")  +
    geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
#plot the intake type
 petpoint%>% ggplot(mapping=aes(x=intake_type,fill=intake_type)) + 
    geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Intake Type")  +
    geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
 #plot the intake sub-type
 petpoint%>% ggplot(mapping=aes(x=intake_subtype,fill=intake_subtype)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   labs(title = "Intake Sub Type") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
 #plot the intake reason
 petpoint%>% ggplot(mapping=aes(x=intake_reason,fill=intake_reason)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(title = "Intake Reason") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 table(petpoint$intake_reason)
 
 #plot the intake sitename
 petpoint%>% ggplot(mapping=aes(x=intake_sitename,fill=intake_sitename)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Intake Sitename") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
  #plot the outcome sitename
 petpoint%>% ggplot(mapping=aes(x=outcome_sitename,fill=outcome_sitename)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Outcome Sitename") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

 #plot the outcome type
 petpoint%>% ggplot(mapping=aes(x=outcome_type,fill=outcome_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Outcome Type") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
  #plot the outcome sub-type
 petpoint%>% ggplot(mapping=aes(x=outcome_subtype,fill=outcome_subtype)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Outcome Sub Type") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
kable(sort(table(petpoint$outcome_subtype),decreasing=TRUE))
 
#How many were altered?
 petpoint%>% ggplot(mapping=aes(x=altered,fill=altered)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Sprayed/Neutered") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
 # How many days between intake and outcome 
petpoint <- petpoint %>% mutate(process_time = interval(intake_date,outcome_date) / ddays(1))
#plot the distribution
petpoint %>% ggplot(mapping=aes(x=process_time)) +
  geom_histogram(breaks=seq(20,300,by=20),col="red",fill="green",alpha=0.2) +
  labs(title="Histogram of Processing Time in Days (Intake -> Outcome)")

#create time period categories
petpoint$process_time_periods <- cut(petpoint$process_time,
      breaks=c(-Inf,1,3,5,10,30,90,180,Inf),
      labels=c("< 1day","2-3 days","4-5 days","6-10 days", "11-30 days", "31-90 days", "91-180days", ">180 days"))
#plot the categories
 ggplot(petpoint,mapping=aes(x=process_time_periods,fill=process_time_periods)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Process Time Periods") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

 
```



