---
title: "PAWS - Actions"
author: "Ramaa Nathan"
date: "3/8/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = "html") 
library(tidyverse)
library(lubridate)
library(skimr)
library(gtools)
library(knitr)  #for outputting tables as kables
```

#### Data Columns:
1. type: type of update done to the card. “createCard” (when the card was created) / “updateCheckItemStateOnCard” (when a checklist item was modified). PAWS use a fairly standardized checklist for going through the applications, so the timeline in which that checklist gets completed can be informative
2. data.checkItem.state: if a checklist item was modified, whether it was modified to “complete” or “incomplete”
3. date: the date when the action was taken
4. data.card.id:
5. animal_type_: cat or dog
6. checklist_ACCT: Check with ACCT (Animal Care and Control Team)
7. checklist_CHQ: Check with clinic HQ (the PAWS clinic uses a different software for logging clients than the adoptions process)
8. checklist_LL: Check with property owner
9. checklist_PP: Check with PetPoint
10. checklist_SPCA: Check with SPCA (Society for the Prevention of Cruelty to Animals)
11. checklist_TR: Check with Trello
12. checklist_VET: Check with vet


```{r cards}
cat_actions = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/cat_actions.csv")
dog_actions = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/dog_actions.csv")
cat_actions
dog_actions

#create one dataframe
actions <- bind_rows(unique(cat_actions),unique(dog_actions))
actions <- actions  %>%
  mutate (animal_type=factor(animal_type))
str(actions)

# Are there any duplicates in the data?
anyDuplicated(actions)

# find number of missing data in each column
# print("Missing Entries")
# skim(actions) %>%
#   filter(stat == "missing") %>% filter(value > 0) 
# actions %>% group_by(animal_type) %>% skim() %>% filter(stat == "missing") %>% filter(value > 0) 
# table(actions$data.checkItem.state)
```
Observations: 
1. There are no duplicates and there are 6,892 observations with missing data only for data.checkItem.state (4,998 for cats and 1,894 for dogs).
2. There are 31,217 records with complete for data.checkitem.state and 870 records with incomplete.

```{r}
# frequency of type
actions %>% count(type)
```
Observations:
1. 6,892 entries of createCard and 32,087 enrties of updateCheckItemStateOnCard

#### Check Range of Date
```{r}
# range of date
actions %>% skim(date)
```
```{r}
#data.card.id - find number of unique entries
n_distinct(actions$data.card.id)
actions %>% group_by(animal_type) %>% summarize(unique_ids = n_distinct(data.card.id))
```
Observations: There are 7,209 unique data,.card.id


```{r}
anyDuplicated(actions)
actions_wide <-
    actions %>% distinct() %>% 
    rename(id = data.card.id) %>%
    gather(item, result, checklist_ACCT:checklist_VET) %>%
    group_by(id) %>% 
    mutate(date_start = min(date)) %>% 
    filter(result == TRUE) %>%
    group_by(id, item) %>%
    filter(date == max(date)) %>%
    ungroup() %>%
    mutate(wait = difftime(date, date_start, units = "days"),
           wait = round(as.numeric(wait), 2)) %>%
    select(-c(date, data.checkItem.state, type, result)) %>%
    distinct() %>%
    spread(item, wait) %>% 
  mutate(wday_start = wday(date_start, label = TRUE, abbr = TRUE))
# 
# actions_wide
#write.csv(actions_wide, "./Data/actions_wide.csv", row.names = FALSE)
anyDuplicated(actions_wide)
actions_wide %>% sample_n(10)
actions %>% filter(data.card.id=="5b12f48abd07fe5382780044")
actions_wide %>% filter(id=="5b12f48abd07fe5382780044")
```
Observations: We get duplicates after computing the wait time as there seem to be multiple entries for some of the checklist item on the same days for a few record -  one with data.checkItem.state==Complete and theother with incomplete state


