---
title: "PAWS - Cats and Dogs Apps"
author: "Ramaa Nathan"
date: "3/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = "html") 
library(tidyverse)
library(lubridate)
library(skimr)
library(gtools)
library(knitr)  #for outputting tables as kables
```
#### Data Transformations
The cat_apps.csv and dog_apps.csv have exactly same columns but there is a mismatch in their datatypes.
So, first step is to fix them
1. Add animal_type="cats" to cats.csv and animal_type="dogs" to dogs.csv 
2. cat_apps.csv - convert adults_in _home from character type to integer type
2. dog_apps.csv - convert zip code from integer type to character type so that we can categorize them


```{r}
cat_apps = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/cat_apps.csv")
dog_apps = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/dog_apps.csv")

#convert adults_in_home to integer type
cat_apps <- cat_apps[,-1] %>% 
  distinct() %>%
  transform(adults_in_home = as.numeric(adults_in_home)) %>%
  mutate(animal_type="cats",
         ZIP=ifelse(str_length(ZIP)<5,str_c("0",ZIP),ZIP))

#convert ZIP to character type and pad it correctly
dog_apps <- dog_apps[,-1] %>% 
  distinct() %>% 
  transform(ZIP = as.character(ZIP)) %>%
  mutate(animal_type="dogs",
         ZIP=ifelse(str_length(ZIP)<5,str_c("0",ZIP),ZIP))          


apps <- bind_rows(cat_apps,dog_apps)

#check for unique values
apps %>% select(orig=reason_for_adoption) %>% unnest(orig=str_split(orig,",")) %>% unique() 
apps %>% select(orig=allergies) %>% unnest(orig=str_split(orig,",")) %>% unique()
apps %>% select(orig=ideal_adoption_timeline) %>% unnest(orig=str_split(orig,",")) %>% unique()
apps %>% select(orig=home_pet_policy) %>% unnest(orig=str_split(orig,",")) %>% unique()
apps %>% select(orig=experience) %>% unnest(orig=str_split(orig,",")) %>% unique()

clean_adoption_timeline <- function(x) {
  x %>% str_replace_all(.,"next-few-weeks","few-weeks")
}

clean_household_agree <- function(x) {
  x %>% str_replace_all(.,"it-s-a-surprise","a-surprise")
}

clean_pet_policy <- function(x) {
  x %>% str_replace_all(.,"no-yet","not-yet") %>%
    str_replace_all(.,"havent-asked","not-yet") %>%
    str_replace_all(.,"n-a","not-applicable")
}

# clean the comma-separated fields
clean_experience <- function(x){
    x %>%
        str_replace_all(., "(grew-up-with)(-pet)", "\\1") %>% 
        #str_replace_all(., "(currently)(-pet)", "\\1") %>% 
        str_replace_all(., "(euthanized)[^,]+", "\\1") %>% 
        str_replace_all(., "had-pet-die", "pet-died-in-care") %>% 
        str_replace_all(., "[^,]*(lived-with-housemate|lived-wit-prev)[^,]*", "past-housemates-pet") %>% 
        str_replace_all(., "currently-pets[^,]*", "current-housemates-pet") %>% 
        str_replace_all(., "(never-lived-)[^,]+", "\\1with") %>% 
        str_replace_all(., "(given-)[^,]*shelter", "\\1to-shelter") %>% 
        str_replace_all(., "(given-)(pet-)?(to-another)", "\\1away") %>% 
        str_replace_all(., "(bred-sold)-a-pet", "bred-sold")
}

clean_budget <- function(x) {
  x %>% str_replace_all(.,"^-","") %>%
    parse_number(.)
}

clean_pet_kept <- function(x) {
  x %>% str_replace_all(.,"unsupervised-access-to-my-yard-9doggie-door-etc","unsupervised-access-to-my-yard-doggie-door-etc")
}

compare_fields <- function(df, x, fun){
    df %>% 
        select(orig = x) %>% 
        filter(!is.na(orig)) %>% 
        unnest(orig = str_split(orig, ",")) %>%
        distinct() %>% 
        mutate(clean = do.call(fun, list(orig))) %>% 
        arrange(clean) %>% 
        mutate(match = ifelse(clean != orig, "*", ""))
}

compare_fields(apps, "experience", clean_experience) %>% filter(match == "*")
#str(apps)
apps <- apps %>% 
  select(-c(STATEFP,COUNTYFP,TRACTCE,GEOID,NAME,NAMELSAD,MTFCC,FUNCSTAT,ALAND,AWATER,INTPTLAT,INTPTLON)) %>%
  rename(trello_id = outcome_trello_id) %>%
  mutate(date_submitted = mdy(date_submitted),
         ideal_adoption_timeline = clean_adoption_timeline(ideal_adoption_timeline),
         all_household_agree = clean_household_agree(all_household_agree),
         home_pet_policy = clean_pet_policy(home_pet_policy),
         experience = clean_experience(experience),
         pet_kept = clean_pet_kept(pet_kept),
         adults_in_home = abs(adults_in_home),
         children_in_home = abs(children_in_home),
         home_alone_avg = parse_number(home_alone_avg),
         home_alone_max = parse_number(home_alone_max),
         budget_monthly = clean_budget(budget_monthly),
         budget_emergency = clean_budget(budget_emergency),
         budget_monthly_ranges = factor(case_when(budget_monthly <=25 ~ "Less than $25",
                                             budget_monthly <=100 ~ "$25-$100",
                                             budget_monthly <=200 ~ "$100-$200",
                                             budget_monthly <=500 ~ "$200-$500",
                                             budget_monthly <=1000 ~ "$500-$1000",
                                             budget_monthly <=5000 ~ "$1000-$5000",
                                             is.na(budget_monthly) ~ "NA",
                                             TRUE ~ ">$5000"),
                                   levels=c("<$25","$25-$100","$100-$200","$200-$500","$500-$1000","$1000-$5000",">$5000","NA"),
                                   ordered=T),
         budget_emergency_ranges = factor(case_when(budget_emergency <=25 ~ "Less than $25",
                                             budget_emergency  <=100 ~ "$25-$100",
                                             budget_emergency  <=200 ~ "$100-$200",
                                             budget_emergency  <=500 ~ "$200-$500",
                                             budget_emergency  <=1000 ~ "$500-$1000",
                                             budget_emergency  <=5000 ~ "$1000-$5000",
                                             is.na(budget_emergency ) ~ "NA",
                                             TRUE ~ ">$5000"),
                                   levels=c("<$25","$25-$100","$100-$200","$200-$500","$500-$1000","$1000-$5000",">$5000","NA"),
                                   ordered=T))

apps %>% ggplot(mapping=aes(x=budget_monthly_ranges,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Monthly budgets") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
apps %>% ggplot(mapping=aes(x=budget_emergency_ranges,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Emergency budgets") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Home alone avg
apps %>% ggplot(mapping=aes(x=home_alone_avg,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Home Alone Average") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
### Home alone max
apps %>% ggplot(mapping=aes(x=home_alone_max,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Home Alone Maximum") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
str(apps)

#apps %>% sample_n(30) %>% select(budget_monthly,budget_monthly_ranges)
#apps %>% select(reason_for_adoption) %>% unnest(rfa=str_split(reason_for_adoption,","))
#convert all other character types to factor types
#mutate_if(sapply(apps,is.character),as.factor)
#str(apps)
#summary(apps)
## CONVERT HOME_ALONE to integer after taking out h....
```

```{r}

# create separate dataframe with indicator columns - one for each option of the provided categorical field
convert_to_ind <- function(df, field){
    df %>% 
        mutate_(var = field) %>% 
        distinct(trello_id, var) %>% 
        unnest(split = str_split(var, ",")) %>%
        select(-var) %>% 
        filter(!is.na(split)) %>% 
        mutate(n = 1,
               split = 
                   str_replace_all(split, "-", ".") %>% 
                   paste0(str_replace_all(field, "_", "."), 
                          "_", ., "_ind")) %>%
        distinct() %>% 
        spread(split, n, fill = 0)
}
 convert_to_ind(apps,"reason_for_adoption") %>% head() %>% View()
 
 str(apps_with_indicators)
```


