---
title: "PAWS - Actions"
author: "Ramaa Nathan"
date: "3/8/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = "html") 
library(tidyverse)
library(lubridate)
library(skimr)
library(gtools)
library(knitr)  #for outputting tables as kables
```

#### Data Columns:
1. type: type of update done to the card. “createCard” (when the card was created) / “updateCheckItemStateOnCard” (when a checklist item was modified). PAWS use a fairly standardized checklist for going through the applications, so the timeline in which that checklist gets completed can be informative
2. data.checkItem.state: if a checklist item was modified, whether it was modified to “complete” or “incomplete”
3. date: the date when the action was taken
4. data.card.id:
5. animal_type_: cat or dog
6. checklist_ACCT: Check with ACCT (Animal Care and Control Team)
7. checklist_CHQ: Check with clinic HQ (the PAWS clinic uses a different software for logging clients than the adoptions process)
8. checklist_LL: Check with property owner
9. checklist_PP: Check with PetPoint
10. checklist_SPCA: Check with SPCA (Society for the Prevention of Cruelty to Animals)
11. checklist_TR: Check with Trello
12. checklist_VET: Check with vet
13. checklist_seq_num: indicates the number of the action on a given card, in a sequential order by date. “createCard” should always be the first entry; if it isn’t, the entry isn’t complete and we should not use it

```{r cards}
cat_actions = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/cat_actions.csv")
dog_actions = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/dog_actions.csv")
cat_actions
dog_actions

#create one dataframe
actions <- bind_rows(cat_actions,dog_actions)
str(actions)

# Are there any duplicates in the data?
anyDuplicated(actions)

# find number of missing data in each column
print("Missing Entries")
skim(actions) %>%
  filter(stat == "missing") %>% filter(value > 0) 
actions %>% group_by(animal_type) %>% skim() %>% filter(stat == "missing") %>% filter(value > 0) 
table(actions$data.checkItem.state)
```
Observations: 
1. There are no duplicates and there are 6,892 observations with missing data only for data.checkItem.state (4,998 for cats and 1,894 for dogs).
2. There are 31,217 records with complete for data.checkitem.state and 870 records with incomplete.

```{r}
# frequency of type
actions %>% count(type)
```
Observations:
1. 6,892 entries of createCard and 32,087 enrties of updateCheckItemStateOnCard

#### Check Range of Date
```{r}
# range of date
actions %>% skim(date)
```
```{r}
#data.card.id - find number of unique entries
n_distinct(actions$data.card.id)
actions %>% group_by(animal_type) %>% summarize(unique_ids = n_distinct(data.card.id))
```
Observations: There are 7,209 unique data,.card.id

#### Remove entries of those card.data.id that do not have createCard as the first entry in their sequence
```{r}
#checklist_seq_num: indicates the number of the action on a given card, in a sequential order by date. “createCard” should always be the first entry; if it isn’t, the entry isn’t complete and we should not use it
#identify the ids whose first entry is not createCard
invalidCardIds <- actions %>% filter(checklist_seq_num==1 & type != "createCard") %>% select(data.card.id) %>% distinct()
invalidCardIds
# filter out the invalid cards ids from teh actions dataset
actions_mod <- actions %>% anti_join(invalidCardIds)
str(actions_mod)
#check if the removal of these records caused a change in the number of missing data cases.
actions_mod %>% group_by(animal_type) %>% skim() %>% filter(stat == "missing") %>% filter(value > 0) 
```
Observations:
1. There are 317 data.card.ids that do note have createCard as the first item in the sequence of updates
2. The removal of the invalidd data.card.ids does not change the number of missing data items.

```{r}
actions_mod %>% filter(checklist_seq_num==2) %>% select(starts_with("checklist")) 
#What is the last seqnum?
actions_mod %>% group_by(data.card.id) %>% 
  summarize(lastSeqNum = max(checklist_seq_num)) %>% 
  ggplot(mapping=aes(x=lastSeqNum)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   labs(title = "Distribution of Number of Actions for the Applications ") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
actions_mod %>% group_by(data.card.id) %>% 
  summarize(lastSeqNum = max(checklist_seq_num)) %>% select(lastSeqNum) %>% table()
actions_mod %>% group_by(data.card.id) %>% 
  summarize(lastSeqNum = max(checklist_seq_num)) %>%  count(lastSeqNum)

# Check the patterns of date and checklist_seq_num grouped by data_card_id
actions_mod %>% group_by(data.card.id) %>% 
  select(data.card.id,date,starts_with("checklist")) %>%
  arrange(data.card.id,checklist_seq_num)
```
Observations:

1. For any data.card.id,  the checklist_seq_num contiguous? Many of them start from 1, then jump to 12,14,15,16. (Example: 5a219daee835d40ffb1adc73)

2. A few others (Example: 5a3977f2860604c06c11460b) have 1 to 7 and then continue with 12, 14,15,16, etc.

3. When we sort by checklist_seq_num within a data_card_id, the data is not monotonically increasing(Example: 5a3977f2860604c06c11460b

4. Basically, many actions seem to be missing.

