---
title: "PAWS - CARDS"
author: "Ramaa Nathan"
date: "3/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = "html") 
library(tidyverse)
library(lubridate)
library(skimr)
library(gtools)
library(knitr)  #for outputting tables as kables
```

#### Data Fields
1. id: a unique ID
2. dateLastActivity: the last time an activity was recorded on the card
3. dueComplete: indicates whether the application is marked as complete or they lost contact with the applicant
4. due: indicates the date the application is marked as complete or they lost contact with the applicant animal_type: cat or dog
5. label_names: the label names indicate the last recorded status of the application


```{r cards}
cat_cards = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/cat_cards.csv")
dog_cards = read_csv("/Users/ramaa/Documents/Meetups/R/PAWS/2019_datathon/Data/dog_cards.csv")
# Add a new column "animal_type" to each dataset
cat_cards <- cat_cards %>% mutate(animal_type="cats");
dog_cards <- dog_cards %>% mutate(animal_type="dogs");
str(cat_cards)
str(dog_cards)

#create one dataframe
cards <- bind_rows(cat_cards,dog_cards)
str(cards)
cards <- cards %>% mutate (animal_type=factor(animal_type))

#check for duplicates
# Are there any duplicates in the data?
anyDuplicated(cards)

# find number of missing data in each column
print("Missing Entries")
cards %>% group_by(animal_type) %>% skim() %>% 
  filter(stat == "missing") %>% filter(value > 0) 
#find percentage of missing entries
cards %>% group_by(animal_type) %>% summarise(n=n(), 
                                              missing_due=sum(is.na(due)),missing_due_percent = 100*missing_due/n,
                                            missing_labels=sum(is.na(label_names)),missing_labels_percent = 100*missing_labels/n)

## Check for duplicate ids
numNonUniquesIDs = cards %>% 
   filter(!is.na(id)) %>% 
   select(id) %>% 
   group_by(id) %>% 
   summarise(dupcount=n()) %>%
  filter(dupcount > 1)

#plot the due Complete by  animal type
 # cards%>% ggplot(mapping=aes(x=dueComplete,col=animal_type,fill=animal_type)) + 
 #     geom_bar() +
 #     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
 #      labs(title = "Due Complete by Animal Type")  +
 #    geom_text(aes(label=..count..),stat='count',position=position_stack(1.1),col="black")


```

#### Extract the last label from label names 
```{r}
extractFromString = function(str1) {
  sl = str_split(str1,",")
  l = length(sl[[1]])
  last = str_trim(sl[[1]][l])
  return (unlist(last)) #(list(l,last))
}
cards$lastLabel = sapply(cards$label_names, FUN=function(x) extractFromString(x))
cards$numLabels = sapply(cards$label_names, FUN=function(x) ifelse(is.na(x),0,length(str_split(x,",")[[1]])))
str(cards)
cards %>% skim(numLabels)
```
#### What is the distribution of labels for each type of dueComplete status?
```{r}
 
#what are the last labels associated with dueComplete=FALSE
cards %>% filter(!is.na(lastLabel) & dueComplete == FALSE) %>% 
  ggplot(mapping=aes(x=lastLabel,fill=lastLabel)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Distribution of lastLabel when dueComplete = FALSE") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
cards %>% filter(dueComplete == FALSE) %>% select(lastLabel) %>% table(useNA="always") %>% sort(decreasing = T)

#what are the last labels associated with dueComplete=TRUE
cards %>% filter(dueComplete) %>% 
  ggplot(mapping=aes(x=lastLabel,fill=lastLabel)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Distribution of lastLabel when dueComplete = TRUE") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
cards %>% filter(dueComplete) %>% select(lastLabel) %>% table(useNA="always") %>% sort(decreasing = T)

#Distribution of lastLabel 
cards %>% filter(!is.na(label_names)) %>% 
  ggplot(mapping=aes(x=lastLabel,fill=lastLabel)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Distribution of lastLabel") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
cards %>% select(lastLabel) %>% table(useNA="always") %>% sort(decreasing = T)

```
Observation: dueComplete=FALSE even when the last status of the applications ends in adoption (712 cases) or ready to adopt(981). 
Observation: dueComplete=TRUE even when the application status is one of  denied(52), red flag(55), etc. 
Also, there are 865 cases when dueComplete is TRUE but there are no labels. So, dueComplete does not seem to be getting updated reliably.

#### Compare Dates
dateLastActivity: the last time an activity was recorded on the card
due: indicates the date the application is marked as complete or they lost contact with the applicant
Basically, we expect dateLastActivity to be always later than or equal to the due date

```{r}
 # Compare dateLastActivity and due
#When a process is completed leading to adoption, the due date should correspond to when the records were marked as complete
#check records when dueComplete is true
#dueComplete: indicates whether the application is marked as complete or they lost contact with the applicant
cards %>% filter(dueComplete & due>dateLastActivity) %>% select(due,dateLastActivity,lastLabel)
cards %>% filter(!dueComplete & due>dateLastActivity) %>% select(due,dateLastActivity,lastLabel)
```
Observation: There are 7 cases when dueComplete is TRUE and due date is later then dateLastActivity. There are 51 cases where dueComplete is FALSE and the due date is later than dateLastActivity

#### Distribution of number of labels by type of last label
```{r}
cards %>% filter(numLabels==6)%>% select(lastLabel) %>% table() %>% sort(decreasing = T)
cards %>% filter(numLabels==5)%>% select(lastLabel) %>% table() %>% sort(decreasing = T) %>% head(3)
cards %>% filter(lastLabel=="adopted")%>% select(numLabels) %>% table() %>% sort(decreasing = T) %>% head(3)
```
Observations: 
1. Number of labels range from 0 to 6. 
2. There are 5 instances when label=6 and the corresponding last labels are: denied, manager decision, pet policy, questions, red flag
3. There are 13 cases with nuber of labels =5 and the corresponding last albels are red flag, adopted elsewhere, and denied
4. When the last label is adopted, the number of labels are either 1 (918 cases), 2 (198), or 3(18)

