---
title: "Initial data exploration"
author: "Amy Goodwin Davies"
always_allow_html: yes
output:
  github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(echo = T, message = F, warning = F, root.dir = "../../../")
```


# Our questions:

* 2.1 How long does each step of an application take? (by animal species and PAWS location where available); i.e., from date of processing start, through checking off each checklist item, to the date and time when an animal was adopted.
* 2.2 What are the applicant and animal characteristics that predict a speedy adoption?
* 2.3 What are the applicant and animal characteristics that predict a denied application?
* 2.4 Are there seasonal patterns to adoptions?
* 2.5 What predicts an applicant who adopts elsewhere?

Load packages

```{r load_packages, echo = T,  message = F,  warning = T}
library(tidyverse)
library(lubridate)
library(data.table)
library(forcats)
library(plyr)
```

Source helper function(s)

```{r source_functions, echo = T,  message = F,  warning = T}
source("Analyses/2_Applicants/helper_functions.R")
```

Load data

```{r load_data, echo = T,  message = F,  warning = T}
cat_actions <- read_csv("Data/cat_actions.csv")
dog_actions <- read_csv("Data/dog_actions.csv")
cat_cards <- read_csv("Data/cat_cards.csv")
dog_cards <- read_csv("Data/dog_cards.csv")
cat_apps <- read_csv("Data/cat_apps.csv")
dog_apps <- read_csv("Data/dog_apps.csv")
petpoint <- read_csv("Data/petpoint.csv")
```

# Actions

Create checklist_names as a factor for each checklist name combination...

```{r actions, echo = T,  message = F,  warning = T}
actions <- rbind(cat_actions, dog_actions)
actions$checklist_names <- ""
actions$checklist_names <- as.character(actions$checklist_names)
actions[actions$checklist_ACCT == TRUE,]$checklist_names <- 
  paste(actions[actions$checklist_ACCT == TRUE,]$checklist_names, "ACCT", sep = "/")
actions[actions$checklist_CHQ == TRUE,]$checklist_names <- 
  paste(actions[actions$checklist_CHQ == TRUE,]$checklist_names, "CHQ", sep = "/")
actions[actions$checklist_LL == TRUE,]$checklist_names <- 
  paste(actions[actions$checklist_LL == TRUE,]$checklist_names, "LL", sep = "/")
actions[actions$checklist_PP == TRUE,]$checklist_names <- 
  paste(actions[actions$checklist_PP == TRUE,]$checklist_names, "PP", sep = "/")
actions[actions$checklist_SPCA == TRUE,]$checklist_names <- 
  paste(actions[actions$checklist_SPCA == TRUE,]$checklist_names, "SPCA", sep = "/")
actions[actions$checklist_TR == TRUE,]$checklist_names <- 
  paste(actions[actions$checklist_TR == TRUE,]$checklist_names, "TR", sep = "/")
actions[actions$checklist_VET == TRUE,]$checklist_names <- 
  paste(actions[actions$checklist_VET == TRUE,]$checklist_names, "VET", sep = "/")
actions[actions$checklist_ACCT == FALSE &
          actions$checklist_CHQ == FALSE &
          actions$checklist_LL == FALSE &
          actions$checklist_PP == FALSE &
          actions$checklist_SPCA == FALSE &
          actions$checklist_TR == FALSE &
          actions$checklist_VET == FALSE,]$checklist_names <- "OTHER"
actions$checklist_names <- trimws(gsub("^/", " ", actions$checklist_names))
actions$checklist_names <- as.factor(actions$checklist_names)
summary(actions$checklist_names)
actions_summary <- ddply(actions, .(checklist_names, animal_type), summarise,
                         count = length(checklist_names))
ggplot(actions_summary, aes(x = fct_reorder(checklist_names, count), y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.25, position = "identity", size = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill=FALSE) +
  facet_wrap(~ animal_type)
```

Make wide version of data (each row corresponds to unique data.card.id)...


```{r actions_2, echo = T,  message = F,  warning = T}
actions <- actions %>%
  dplyr::arrange(date) %>%
  dplyr::group_by(data.card.id) %>%
  dplyr::mutate(new_checklist_seq_num = seq_along(date))
wide_actions <- dcast(actions, formula = data.card.id + animal_type ~ new_checklist_seq_num, value.var = c("checklist_names"))
nrow(wide_actions) == length(unique(actions$data.card.id))
wide_actions <- wide_actions %>% mutate_if(is.character, as.factor)
```

# Cards

```{r cards, echo = T,  message = F,  warning = T}
cards <- rbind(dog_cards, cat_cards)
head(cards)
length(unique(cards$id)) == nrow(cards)
```

Create boolean variables for each label name...

```{r cards_2, echo = T,  message = F,  warning = T}
str(cards$label_names)
summary(cards$label_names)
length(unique(cards$label_names))
label_names_tidy <- tidy_elements(cards, "label_names") # slow...
cards <- label_names_tidy$output_df
label_names_elements <- label_names_tidy$elements
label_names_new_colnames <- label_names_tidy$new_colnames
label_names_elements_summary <- label_names_tidy$elements_summary
ggplot(label_names_elements_summary, aes(x = fct_reorder(label_names, count), y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.25, position = "identity", size = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill=FALSE)
label_names_elements
label_names_new_colnames
write_csv(cards, "Analyses/2_Applicants/amygood/output/cards_edited.csv")
```