---
title: "Label names"
author: "Amy Goodwin Davies"
always_allow_html: yes
output:
  github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(echo = T, message = F, warning = F, root.dir = "../../../")
```

```{r packages, echo = T,  message = F,  warning = T}
library(tidyverse)
library(lubridate)
```

```{r source_functions, echo = T,  message = F,  warning = T}
source("Analyses/2_Applicants/helper_functions.R")
```

# Summary of all label names

```{r read, echo = T,  message = F,  warning = T}
cat_cards <- read_csv("Data/cat_cards.csv")
dog_cards <- read_csv("Data/dog_cards.csv")
cards <- rbind(dog_cards, cat_cards)
head(cards)
length(unique(cards$id)) == nrow(cards)
```

```{r tidy_elements, echo = T,  message = F,  warning = T}
str(cards$label_names)
length(unique(cards$label_names))
label_names_tidy <- tidy_elements(cards, "label_names")
cards <- label_names_tidy$output_df
label_names_elements <- label_names_tidy$elements
label_names_new_colnames <- label_names_tidy$new_colnames
label_names_elements_summary <- label_names_tidy$elements_summary
ggplot(label_names_elements_summary, aes(x = fct_reorder(label_names, count), y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.25, position = "identity", size = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill=FALSE)
```

```{r check_function, echo = T,  message = F,  warning = T}
label_names_new_colnames
all(grepl("ready for review", cards$label_names) == cards$label_names_ready_for_review, na.rm = TRUE)
all(grepl("denied", cards$label_names) == cards$label_names_denied, na.rm = TRUE)
all(grepl("vet check in process", cards$label_names) == cards$label_names_vet_check_in_process, na.rm = TRUE)
all(grepl("vet", cards$label_names) == cards$label_names_vet, na.rm = TRUE) # should be false as "vet" is a substring of other label names
```


```{r visualise_elements, echo = T,  message = F,  warning = T}
# adapted from: https://www.kaggle.com/ranliu/r-ladies-philly-intro-to-machine-learning-meetup
cards$animal_type <- as.factor(cards$animal_type)

true_by_column <- cards %>% 
    select(label_names_new_colnames, animal_type) %>%
    as_data_frame() %>%
    mutate(row_number = 1:nrow(.)) %>%
    gather(key = variable, value = is_true, -row_number, -animal_type) 

ggplot(true_by_column, aes(y = variable, x = row_number,
                           fill = is_true)) +
    geom_tile() +
    facet_grid(. ~ animal_type, scales = "free") +
    scale_fill_grey() +
    labs(y = "Indicator variables",
         x = "Rows")

table(cards$animal_type, is.na(cards$label_names))
```


* use ontology from [codebook](../../../Data/data_codebook.md) to create subgroups of label names


## Label names relating to application approval status

* *approved*
* *ready to adopt*
* *ready for review* - means ready to adopt but applicant needs to be given additional info to set up for success
* *reviewed with handouts only* - ready to adopt, should be given a handout
* *approved with limitation*
* *needs review before approval*
* *denied*
* *withdrawn*
* *do not follow up*
* *manager decision*

```{r approval, echo = T,  message = F,  warning = T}
approval_indicators <- c("label_names_approved",
                        "label_names_ready_to_adopt",
                        "label_names_ready_for_review",
                        "label_names_reviewed_with_handouts_only",
                        "label_names_approved_with_limitation",
                        "label_names_needs_review_before_approval",
                        "label_names_denied",
                        "label_names_withdrawn",
                        "label_names_do_not_follow_up",
                        "label_names_manager_decision")
true_by_column_approval_indicators <- cards %>% 
    select(approval_indicators, animal_type) %>%
    as_data_frame() %>%
    mutate(row_number = 1:nrow(.)) %>%
    gather(key = variable, value = is_true, -row_number, -animal_type) 

summary(true_by_column_approval_indicators)

ggplot(true_by_column_approval_indicators, aes(y = variable, x = row_number,
                           fill = is_true)) +
    geom_tile() +
    facet_grid(. ~ animal_type, scales = "free") +
    scale_fill_grey() +
    labs(y = "Indicator variables",
         x = "Rows")

approval_indicators_summary <- cards %>%
    group_by(animal_type) %>% 
    summarise_at(approval_indicators, sum, na.rm = TRUE) %>%
    gather(approval_indicators, count, -animal_type)

ggplot(approval_indicators_summary, aes(x = fct_reorder(approval_indicators, count), y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.25, position = "identity", size = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(. ~ animal_type) +
  guides(fill=FALSE)
```

## Label names relating to issues with the application

* *vet*
* *vet check in process*
* *need vet info*
* *need roommates vet info*
* *not utd* - applicant’s current animals are not up to date on vaccines
* *questions* - follow up questions about pets they had before without explanation
* *not s n* - KF thinks that this means applicant’s current animals are not spayed or neutered
* *dog meet* - dog meet needs to occur before proceeding
* *need to see id* - applicant didn’t provide their ID info on application
* *need info* - need additional info
* *checks* - checklist needs to be added and items gone through
* *needs app attached*
* *rescue check* - check applicant’s name with other rescues
* *pet policy* - pet policy needs to be confirmed
* *need written ll permission* - need written landlord permission
* *opa* - owner of property where applicant lives must be confirmed
* *need proof of ownership* - applicant said they own their property but this info cannot be found publicly
* *landlord* - must check with landlord
* label relating to limitations of the application:
* *declaw only* - will only adopt a declawed cat
* *red flag* - some issue with applicant, might not be enough to disqualify but could require follow up
* *unsure foster or adopt* - applicant might be ambivalent about fostering or adopting
* *foster to adopt* - applicant will foster an animal until the animal can be fixed, when it is legal for PAWS to adopt out animal
* *serial no show*
```{r issues, echo = T,  message = F,  warning = T}
issues_indicators <- c("label_names_vet",
                                 "label_names_vet_check_in_process",
                                 "label_names_need_vet_info",
                                 "label_names_need_roommates_vet_info",
                                 "label_names_not_utd",
                                 "label_names_questions",
                                 "label_names_not_s_n",
                                 "label_names_dog_meet",
                                 "label_names_need_to_see_id",
                                 "label_names_need_info",
                                 "label_names_checks",
                                 "label_names_needs_app_attached",
                                 "label_names_rescue_check",
                                 "label_names_pet_policy",
                                 "label_names_need_written_ll_permission",
                                 "label_names_opa",
                                 "label_names_need_proof_of_ownership",
                                 "label_names_landlord")
true_by_column_issues_indicators <- cards %>% 
    select(issues_indicators, animal_type) %>%
    as_data_frame() %>%
    mutate(row_number = 1:nrow(.)) %>%
    gather(key = variable, value = is_true, -row_number, -animal_type) 

summary(true_by_column_issues_indicators)

ggplot(true_by_column_issues_indicators, aes(y = variable, x = row_number,
                           fill = is_true)) +
    geom_tile() +
    facet_grid(. ~ animal_type, scales = "free") +
    scale_fill_grey() +
    labs(y = "Indicator variables",
         x = "Rows")

issues_indicators_summary <- cards %>%
    group_by(animal_type) %>% 
    summarise_at(issues_indicators, sum, na.rm = TRUE) %>%
    gather(issues_indicators, count, -animal_type)

ggplot(issues_indicators_summary, aes(x = fct_reorder(issues_indicators, count), y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.25, position = "identity", size = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(. ~ animal_type) +
  guides(fill=FALSE)
```

## Label names relating to limitations of the application

* *declaw only* - will only adopt a declawed cat
* *red flag* - some issue with applicant, might not be enough to disqualify but could require follow up
* *unsure foster or adopt* - applicant might be ambivalent about fostering or adopting
* *foster to adopt* - applicant will foster an animal until the animal can be fixed, when it is legal for PAWS to adopt out animal
* *serial no show*

```{r limitations, echo = T,  message = F,  warning = T}
limitations_indicators <- c("label_names_declaw_only",
                                 "label_names_red_flag",
                                 "label_names_unsure_foster_or_adopt",
                                 "label_names_foster_to_adopt",
                                 "label_names_serial_no_show")
true_by_column_limitations_indicators <- cards %>% 
    select(limitations_indicators, animal_type) %>%
    as_data_frame() %>%
    mutate(row_number = 1:nrow(.)) %>%
    gather(key = variable, value = is_true, -row_number, -animal_type) 

summary(true_by_column_limitations_indicators)

ggplot(true_by_column_limitations_indicators, aes(y = variable, x = row_number,
                           fill = is_true)) +
    geom_tile() +
    facet_grid(. ~ animal_type, scales = "free") +
    scale_fill_grey() +
    labs(y = "Indicator variables",
         x = "Rows")

limitations_indicators_summary <- cards %>%
    group_by(animal_type) %>% 
    summarise_at(limitations_indicators, sum, na.rm = TRUE) %>%
    gather(limitations_indicators, count, -animal_type)

ggplot(limitations_indicators_summary, aes(x = fct_reorder(limitations_indicators, count), y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.25, position = "identity", size = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(. ~ animal_type) +
  guides(fill=FALSE)
```

## Label names relating to adoption status

* *adopted*
* *adopted elsewhere*
* *adoption follow up* - to be followed up with post-adoption
* *returned*

```{r adoption_status, echo = T,  message = F,  warning = T}
adoption_status_indicators <- c("label_names_adopted",
                                 "label_names_adopted_elsewhere",
                                 "label_names_adoption_follow_up",
                                 "label_names_returned")
true_by_column_adoption_status_indicators <- cards %>% 
    select(adoption_status_indicators, animal_type) %>%
    as_data_frame() %>%
    mutate(row_number = 1:nrow(.)) %>%
    gather(key = variable, value = is_true, -row_number, -animal_type) 

summary(true_by_column_adoption_status_indicators)

ggplot(true_by_column_adoption_status_indicators, aes(y = variable, x = row_number,
                           fill = is_true)) +
    geom_tile() +
    facet_grid(. ~ animal_type, scales = "free") +
    scale_fill_grey() +
    labs(y = "Indicator variables",
         x = "Rows")

adoption_status_indicators_summary <- cards %>%
    group_by(animal_type) %>% 
    summarise_at(adoption_status_indicators, sum, na.rm = TRUE) %>%
    gather(adoption_status_indicators, count, -animal_type)

ggplot(adoption_status_indicators_summary, aes(x = fct_reorder(adoption_status_indicators, count), y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.25, position = "identity", size = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(. ~ animal_type) +
  guides(fill=FALSE)
```

checking...

```{r check, echo = T,  message = F,  warning = T}
all(label_names_new_colnames %in%
        c(approval_indicators, issues_indicators, limitations_indicators, adoption_status_indicators))
```

# To do

* Relationship between cards label_names and actions checklist items
* Positive/negative within each category
