---
title: "combined_dataset"
date: "3/17/2019"
always_allow_html: yes
output: 
  github_document:
    toc: true 
    df_print: kable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F, root.dir = "../../")
```

```{r scipen, echo = T,  message = F,  warning = T}
# Turn off scientific notation ----
options(scipen=999) #digits = 2
```

```{r load_packages, echo = FALSE, message=F, warning=F}
#load packages
library(readr)
library(dplyr)
library(tidyr)

library(tidyverse)
library(lubridate)
library(skimr)
library(gtools)
library(knitr)  #for outputting tables as kables
```

##Load Data
```{r load_data, echo = F,  message = F,  warning = F}
cat_apps = read_csv("../../Data/cat_apps.csv")
dog_apps = read_csv("../../Data/dog_apps.csv")

cat_actions = read_csv("../../Data/cat_actions.csv")
dog_actions = read_csv("../../Data/dog_actions.csv")

petpoint = read_csv("../../Data/petpoint.csv")

cat_cards = read_csv("../../Data/cat_cards.csv")
dog_cards = read_csv("../../Data/dog_cards.csv")
```

##Helper Functions

convert_to_ind(): creates indicator for specified field in dataframe

```{r helper_functions, echo = T,  message = F,  warning = T}
# create separate dataframe with indicator columns - one for each option of the provided categorical field
convert_to_ind <- function(df, field){
    df %>% 
        mutate_(var = field) %>% 
        distinct(trello_id, animal_type, var) %>% 
        unnest(split = str_split(str_trim(var), ",")) %>%
        select(-var) %>% 
        filter(!is.na(split)) %>% 
        mutate(split = str_trim(split)) %>%
        mutate(n = 1,
               split = 
                   str_replace_all(split, "-", ".") %>% 
                   str_replace_all(., " ", ".") %>%
                   paste0(str_replace_all(field, "_", "."), 
                          "_", ., "_ind")) %>%
        distinct() %>% 
        spread(split, n, fill = 0)
}

clean_adoption_timeline <- function(x) {
  x %>% str_replace_all(.,"next-few-weeks","few-weeks")
}

clean_household_agree <- function(x) {
  x %>% str_replace_all(.,"it-s-a-surprise","a-surprise") %>% 
    str_replace_all(.,"yes,no","no,yes") %>%
    str_replace_all(.,"a-surprise,yes","yes,a-surprise")
}

clean_pet_policy <- function(x) {
  x %>% str_replace_all(.,"no-yet","not-yet") %>%
    str_replace_all(.,"havent-asked","not-yet") %>%
    str_replace_all(.,"n-a","not-applicable")
}

# clean the comma-separated fields
clean_experience <- function(x){
    x %>%
        str_replace_all(., "(grew-up-with)(-pet)", "\\1") %>% 
        #str_replace_all(., "(currently)(-pet)", "\\1") %>% 
        str_replace_all(., "(euthanized)[^,]+", "\\1") %>% 
        str_replace_all(., "had-pet-die", "pet-died-in-care") %>% 
        str_replace_all(., "[^,]*(lived-with-housemate|lived-wit-prev)[^,]*", "past-housemates-pet") %>% 
        str_replace_all(., "currently-pets[^,]*", "current-housemates-pet") %>% 
        str_replace_all(., "(never-lived-)[^,]+", "\\1with") %>% 
        str_replace_all(., "(given-)[^,]*shelter", "\\1to-shelter") %>% 
        str_replace_all(., "(given-)(pet-)?(to-another)", "\\1away") %>% 
        str_replace_all(., "(bred-sold)-a-pet", "bred-sold")
}

clean_pet_kept <- function(x) {
  x %>% str_replace_all(.,"unsupervised-access-to-my-yard-9doggie-door-etc","unsupervised-access-to-my-yard-doggie-door-etc")
}

#budget cleanup functions
clean_budget <- function(x) {
  x %>% str_replace_all(.,"^-","") %>%
    parse_number(.) %>%
    gsub("[$]|[(]|[)]|[,]", "", .) %>% 
    as.numeric()
}

create_budget_range <- function(x) {
    case_when( x <= 25 ~ "<25",
               x <= 100 ~ "26-100", 
               x <= 200 ~ "101-200",
               x <= 500 ~ "201-500",
               x <= 1000 ~ "501-1000",
               x <= 5000 ~ "1001-5000",
               is.na(x) ~ "NA",
               TRUE ~ ">5000")
}

#useful functions for QA
get_unique_elements <- function(df, colname) {
  elements_string <- do.call(paste, c(as.list(df[colname]), sep = ","))
  elements_list <- unique(trimws(unlist(strsplit(elements_string, c(",")))))
  unique_elements <- elements_list[!elements_list %in% c("","NA")]
  return(unique_elements)
}

get_elements_summary <- function(output_df, colname, new_colnames) {
  subset_df <- output_df[names(output_df) %in% new_colnames]
  elements_summary <- subset_df %>%
    summarise_all(sum, na.rm = TRUE) %>%
    gather(!!colname, "count")
  return(elements_summary)
}

#city cleanup
clean_city <- function(colname) {
  colname %>% toupper(.) %>%
  gsub("[.]|[,]| PA$", "", .) %>%
  gsub("  ", " ", .) %>%
  gsub("MT ", "MOUNT ", .) %>%
  gsub("19010", "BRYN MAWR", .) %>%
  gsub("CHETSER", "CHESTER", .) %>%
  gsub("ROYERFORD", "ROYERSFORD", .) %>%
  gsub("NORTH WHALES", "NORTH WALES", .) %>%
  gsub("MONTGOMERY VALLAGE", "MONTGOMERY VILLAGE", .) %>%
  gsub("E LANSDOWNE", "EAST LANSDOWNE", .) %>%
  gsub("PHILLY|FILADELFIA|PHILIDELPHIA|PHIMADELPHIA|PHIALADELPHIA|PHIALDELPHIA|PHILDELPHIA", "PHILADELPHIA", .)
}


```

##Dataset Cleanup

### Apps Data

The cat_apps.csv and dog_apps.csv have exactly same columns but there is a mismatch in their datatypes.
<br><br>Steps to fix this:
<br>1. Add animal_type="cat" to cats.csv and animal_type="dog" to dogs.csv 
<br>2. cat_apps.csv - convert adults_in _home from character type to integer type
<br>3. dog_apps.csv - convert zip code from integer type to character type so that we can categorize them

combined dog and cat apps data [1906, 27]

```{r apps, echo = T,  message = F,  warning = T}
#convert adults_in_home to integer type
cat_apps <- cat_apps[,-1] %>% 
  distinct() %>%
  transform(adults_in_home = as.numeric(adults_in_home)) %>%
  mutate(animal_type="cat",
         ZIP=ifelse(str_length(ZIP)<5,str_c("0",ZIP),ZIP))

#convert ZIP to character type and pad it correctly
dog_apps <- dog_apps[,-1] %>% 
  distinct() %>% 
  transform(ZIP = as.character(ZIP)) %>%
  mutate(animal_type="dog",
         ZIP=ifelse(str_length(ZIP)<5,str_c("0",ZIP),ZIP))          

#combine cat and dog apps data
apps <- bind_rows(cat_apps,dog_apps) 
```

Apps data needs to be cleaned up:
      <ul> 
        <li> __Removed columns__: STATEFP,COUNTYFP,TRACTCE,GEOID,NAME,NAMELSAD,MTFCC,FUNCSTAT,ALAND,AWATER,INTPTLAT,INTPTLON </li>          <li> Take absolute value of negative numbers </li>
        <li> Any values > 15 for __adults_in_home__ and __children_in_home__ are updated N/A </li>
        <li> __ideal_adoption_timeline__: consolidated next-few-weeks to few-weeks </li>
        <li> __all_household_agree__: consolidated using the function clean_household_agree </li>
        <li> __home_owner__ and __home_pet_policy__: make factor </li>
        <li> __experience__: cleaned using function above </li>
        <li> __City__: cleaned to reduce same cities with different spellings </li>
        <li> __budget_monthly__ and __budget_emergency__: </li>
<br>          1. Cleaned up the syntax (e.g. () and $)
<br>          2. Updated to take the absolute values (e.g. -3000 is no 3000)
<br>          3. Categorized into buckets (e.g. 100-200)
      </ul>
      

```{r apps_cleanup, echo = T, message = F, warning = T}
apps <- apps %>% 
  select(-c(STATEFP,COUNTYFP,TRACTCE,GEOID,NAME,NAMELSAD,MTFCC,FUNCSTAT,ALAND,AWATER,INTPTLAT,INTPTLON)) %>%
  rename(trello_id = outcome_trello_id) %>%
  mutate(date_submitted = mdy(date_submitted),
         ideal_adoption_timeline = clean_adoption_timeline(ideal_adoption_timeline),
         all_household_agree = clean_household_agree(all_household_agree),
         home_pet_policy = clean_pet_policy(home_pet_policy),
         home_pet_policy = as.factor(home_pet_policy),
         home_owner = as.factor(home_owner),
         experience = clean_experience(experience),
         pet_kept = clean_pet_kept(pet_kept),
         adults_in_home = abs(adults_in_home),
         adults_in_home = replace(adults_in_home, adults_in_home > 15,NA),
         children_in_home = abs(children_in_home), #remove negative numbers
         children_in_home = replace(children_in_home, children_in_home > 15,NA), #remove any numbers greater than 15
         home_alone_avg = parse_number(home_alone_avg),
         home_alone_max = parse_number(home_alone_max),
         budget_monthly = clean_budget(budget_monthly),
         budget_emergency = clean_budget(budget_emergency),
         budget_monthly_ranges = as.factor(create_budget_range(budget_monthly)),
         budget_emergency_ranges = as.factor(create_budget_range(budget_emergency)))
         
#Cleanup city column
apps$City = clean_city(apps$City)
apps$City = replace(apps$City, apps$City %in% c("Y"),NA)
apps$City = as.factor(apps$City)

#Make State factor
apps$State <- as.factor(apps$State)
         
#only extract zip codes with 5 values
apps$ZIP <- str_extract(apps$ZIP, "^.{5}")

apps_with_indicators <- apps %>%
   #distinct(trello_id) %>%
   left_join(convert_to_ind(apps,"reason_for_adoption")) %>%
   left_join(convert_to_ind(apps,"all_household_agree")) %>%
   left_join(convert_to_ind(apps,"allergies")) %>%
   left_join(convert_to_ind(apps,"home_owner")) %>%
   left_join(convert_to_ind(apps,"home_pet_policy")) %>%
   left_join(convert_to_ind(apps,"experience")) %>%
   left_join(convert_to_ind(apps,"budget_monthly_ranges")) %>%
   left_join(convert_to_ind(apps,"budget_emergency_ranges")) %>%
   left_join(convert_to_ind(apps,"home_alone_avg")) %>%
   left_join(convert_to_ind(apps,"home_alone_max")) %>%
   left_join(convert_to_ind(apps,"pet_kept")) %>%
   left_join(convert_to_ind(apps,"exercise")) %>%
   left_join(convert_to_ind(apps,"needs")) %>%
   left_join(convert_to_ind(apps,"return_pet"))


dim(apps)
dim(apps_with_indicators)

colnames(apps_with_indicators)
all(duplicated(apps_with_indicators) == FALSE) # check for duplicates

#dput(get_unique_elements(apps_with_indicators, "experience")) #check unique values
#length(get_unique_elements(apps_with_indicators, "City"))
```

### Actions Dataset
```{r actions, echo = T,  message = F,  warning = T}
#create one dataframe
actions <- bind_rows(unique(cat_actions),unique(dog_actions))

actions <-
    actions %>% distinct() %>% 
    #mutate (animal_type=factor(animal_type)) %>%
    rename(trello_id = data.card.id) %>%
    gather(item, result, checklist_ACCT:checklist_VET) %>%
    group_by(trello_id) %>% 
    mutate(date_start = min(date)) %>% 
    filter(result == TRUE) %>%
    group_by(trello_id, item) %>%
    filter(date == max(date)) %>%
    ungroup() %>%
    mutate(wait = difftime(date, date_start, units = "days"),
           wait = round(as.numeric(wait), 2)) %>%
    select(-c(date, data.checkItem.state, type, result)) %>%
    distinct() %>%
    spread(item, wait) %>% 
    mutate(wday_start = wday(date_start, label = TRUE, abbr = TRUE))
str(actions)
```

### Petpoint Dataset
```{r petpoint, echo = T,  message = F,  warning = T}
petpoint <- petpoint %>% 
  select(-X1) %>% # there are 77 duplicates
  distinct() %>%
  filter(animal_type != "Wildlife") %>%
  select(-c(age_group,STATEFP,COUNTYFP,TRACTCE,GEOID,NAME,NAMELSAD,MTFCC,FUNCSTAT,ALAND,AWATER,INTPTLAT,INTPTLON)) %>%
  rename(trello_id = outcome_trello_id) %>%
  mutate(dob=mdy(dob),
         animal_type=str_to_lower(animal_type),
         intake_date=mdy_hm(intake_date,tz="America/New_York"),
         release_date=mdy_hm(release_date,tz="America/New_York"),
         outcome_date=mdy_hm(outcome_date,tz="America/New_York"),
         outcome_ZIP=as.character(outcome_ZIP),
         outcome_ZIP=ifelse(str_length(outcome_ZIP)<5,str_c("0",outcome_ZIP),outcome_ZIP),
         new_age_group = factor(case_when(age_intake<=1 ~ "<4 weeks",
                                   age_intake <= 3 ~ "4-12 weeks",
                                   age_intake <= 6 ~ "12weeks-6months",
                                   age_intake <= 12 ~ "6months-1year",
                                   age_intake <= 24 ~ "1-2years",
                                   age_intake <= 48 ~ "2-4years",
                                   age_intake <= 72 ~ "4-6years",
                                   age_intake <= 120 ~ "6-10years",
                                   is.na(age_intake) ~ "NA",
                                   TRUE ~ "older than 10years"),
                                levels=c("<4 weeks","4-12 weeks","12weeks-6months",
                                         "6months-1year","1-2years","2-4years",
                                         "4-6years","6-10years","older than 10years","NA"),
                                ordered=TRUE),
         process_time = (interval(intake_date,outcome_date) / ddays(1)),
         process_time_periods = cut(process_time,
                                    breaks=c(-Inf,1,3,5,10,30,90,180,Inf),
                                    labels=c("< 1day","2-3 days","4-5 days","6-10 days", "11-30 days", "31-90 days", "91-180days", ">180 days"))
         ) 

#Spread out the new_group data into different columns
petpoint_with_indicators <- petpoint %>%
  #distinct(trello_id) %>%
  left_join(convert_to_ind(petpoint,"new_age_group"))
    
str(petpoint)

# Some duplicates remain
all(duplicated(petpoint) == FALSE)  
petpoint[duplicated(petpoint),]
petpoint[petpoint$trello_id %in% c("5abd1fc3553a150daabdca1b", "5bd0fef0fbda7d61758333dc"),]
```

### Cards Dataset
```{r cards, echo = T,  message = F,  warning = T}
# Add a new column "animal_type" to each dataset
cat_cards <- cat_cards %>% mutate(animal_type="cat");
dog_cards <- dog_cards %>% mutate(animal_type="dog");
#combine
cards <- bind_rows(cat_cards,dog_cards)

#dueComplete has been found to be unreliable - so remove it
cards <- cards %>% select(-dueComplete) %>%
  rename(trello_id = id) %>%
  mutate (last_label = sapply(cards$label_names, FUN=function(x)
            unlist(
              str_trim(
                tail(
                  str_split(x,",")[[1]],
                  1)))),
          num_labels = sapply(cards$label_names, FUN=function(x)
            ifelse(is.na(x),0,length(str_split(x,",")[[1]]))))

# convert dateLastActivity & due from character to Date
cards <- cards %>%
  mutate(dateLastActivity = mdy(dateLastActivity)) %>%
  mutate(due = mdy(due))

str(cards)

cards_with_indicators <- cards %>%
   #distinct(trello_id) %>%
   left_join(convert_to_ind(cards,"label_names"))

cards_with_indicators %>% sample_n(10) %>% select(contains("label"))
```

```{r}
# find the unique animal_types in each dataset
apps %>% select(animal_type) %>% distinct()
cards %>% select(animal_type) %>% distinct()
actions %>% select(animal_type) %>% distinct()
petpoint %>% select(animal_type) %>% distinct()
```

##Merge the 4 Datasets

```{r}
# master_apps <- apps_with_indicators %>%
#  filter(!is.na(trello_id)) %>%
#  inner_join(petpoint_with_indicators,by = "trello_id") %>%
#  left_join(actions,by = "trello_id") %>%
#  left_join(cards_with_indicators, by = "trello_id")          

master_apps <- apps_with_indicators %>%
 filter(!is.na(trello_id)) %>%
 left_join(actions) %>%
 left_join(cards_with_indicators) %>%
left_join(petpoint_with_indicators)

dim(master_apps)
summary(master_apps)
#write.csv(master_apps, "Analyses/2_Applicants/master_apps.csv", row.names = FALSE)

# Save a single object to a file
datedfilename <- paste("masterapps_",paste(today()) %>% str_replace_all("-",""),".rds",sep="")
saveRDS(master_apps, datedfilename)
# Restore it under a different name to verify
reloaded_data <- readRDS(datedfilename)

str(reloaded_data)
identical(master_apps,reloaded_data)
```

# save processed actions, cards_with_indicators, and petpoint_with_indicators
```{r}
saveRDS(actions, "processed_actions.rds")
saveRDS(cards_with_indicators, "processed_cards_with_indicators.rds")
saveRDS(petpoint_with_indicators, "processed_petpoint_with_indicators.rds")
```

###Merge the Apps and Cards datasets

```{r}
apps_cards <- apps_with_indicators %>%
 filter(!is.na(trello_id)) %>%
 left_join(cards_with_indicators)          

dim(apps_cards)
summary(apps_cards)
#write.csv(master_apps, "Analyses/2_Applicants/master_apps.csv", row.names = FALSE)

# Save a single object to a file
saveRDS(apps_cards, "appsCards.rds")
# Restore it under a different name to verify
#reloaded_data <- readRDS("appsCards.rds")
#str(reloaded_data)
#identical(apps_cards,reloaded_data)
```

## Data Visualizations
### Petpoint Visualizations
```{r,echo = FALSE}
#Distribution of species
 petpoint%>% ggplot(mapping=aes(x=species,fill=species)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Distribution of Species") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

#age group
petpoint %>% ggplot(mapping=aes(x=new_age_group,col=animal_type,fill=animal_type)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Age Group Distribution") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))


 #Distribution of agency name
 petpoint%>% ggplot(mapping=aes(x=agency_name,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Distribution of Agency Name") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
#plot the intake asilomar
 petpoint%>% ggplot(mapping=aes(x=intake_asilomar,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(title = "Intake Asilomar")  +
    geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
#plot the intake type
 petpoint%>% ggplot(mapping=aes(x=intake_type,col=animal_type,fill=animal_type)) + 
    geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Intake Type")  +
    geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
 #plot the intake sub-type
 petpoint%>% ggplot(mapping=aes(x=intake_subtype,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   labs(title = "Intake Sub Type") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
 #plot the intake reason
 petpoint%>% ggplot(mapping=aes(x=intake_reason,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      labs(title = "Intake Reason") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

 
 #plot the intake sitename
 petpoint%>% ggplot(mapping=aes(x=intake_sitename,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Intake Sitename") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
  #plot the outcome sitename
 petpoint%>% ggplot(mapping=aes(x=outcome_sitename,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Outcome Sitename") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

 #plot the outcome type
 petpoint%>% ggplot(mapping=aes(x=outcome_type,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Outcome Type") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
  #plot the outcome sub-type
 petpoint%>% ggplot(mapping=aes(x=outcome_subtype,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Outcome Sub Type") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 
#How many were altered?
 petpoint%>% ggplot(mapping=aes(x=altered,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Sprayed/Neutered") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
 

#plot the distribution of the processing time - number of days between intake and outcome 
petpoint %>% ggplot(mapping=aes(x=process_time,col=animal_type,fill=animal_type)) +
  geom_histogram(breaks=seq(20,300,by=20),col="red",fill="green",alpha=0.2) +
  labs(title="Histogram of Processing Time in Days (Intake -> Outcome)")

#plot the process time periods categories
petpoint %>%  ggplot(mapping=aes(x=process_time_periods,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Process Time Periods") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

```

### Cards Visualizations
```{r}
#Distribution of lastLabel 
cards %>% filter(!is.na(label_names)) %>% 
  ggplot(mapping=aes(x=last_label,col=animal_type,fill=animal_type)) + 
     geom_bar() +
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Distribution of Last Label in the sequence of Label_Names") +
  geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))
cards %>% select(last_label) %>% table(useNA="always") %>% sort(decreasing = T)
```

### Apps Visualizations
```{r}

apps %>% ggplot(mapping=aes(x=budget_monthly_ranges,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Monthly budgets") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

apps %>% ggplot(mapping=aes(x=budget_emergency_ranges,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Emergency budgets") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Home alone avg
apps %>% ggplot(mapping=aes(x=home_alone_avg,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Home Alone Average") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

### Home alone max
apps %>% ggplot(mapping=aes(x=home_alone_max,col=animal_type,fill=animal_type)) +
   geom_bar() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Home Alone Maximum") +
   geom_text(aes(label=..count..),stat='count',position=position_stack(1.1))

```


### Actions Visualizations
```{r}
#plot the distribution of the action time - number of days between first action and checklist_VET 
actions %>% ggplot(mapping=aes(x=checklist_ACCT,col=animal_type,fill=animal_type)) +
  geom_histogram(binwidth=2,col="red",fill="green",alpha=0.2) +
  coord_cartesian(xlim=c(0,25)) +
  facet_grid(animal_type ~ .) +
  labs(title="Histogram of time to action (checklist_ACCT)")


actions %>% ggplot(mapping=aes(x=checklist_CHQ,col=animal_type,fill=animal_type)) +
  geom_histogram(binwidth=2,col="red",fill="green",alpha=0.2) +
  coord_cartesian(xlim=c(0,25)) +
  facet_grid(animal_type ~ .) +
  labs(title="Histogram of time to action (checklist_CHQ)")

actions %>% ggplot(mapping=aes(x=checklist_LL,col=animal_type,fill=animal_type)) +
  geom_histogram(binwidth=2,col="red",fill="green",alpha=0.2) +
  coord_cartesian(xlim=c(0,25)) +
  facet_grid(animal_type ~ .) +
  labs(title="Histogram of time to action (checklist_LL)")

actions %>% ggplot(mapping=aes(x=checklist_PP,col=animal_type,fill=animal_type)) +
  geom_histogram(binwidth=2,col="red",fill="green",alpha=0.2) +
  coord_cartesian(xlim=c(0,25)) +
  facet_grid(animal_type ~ .) +
  labs(title="Histogram of time to action (checklist_PP)")

actions %>% ggplot(mapping=aes(x=checklist_SPCA,col=animal_type,fill=animal_type)) +
  geom_histogram(binwidth=2,col="red",fill="green",alpha=0.2) +
  coord_cartesian(xlim=c(0,25)) +
  facet_grid(animal_type ~ .) +
  labs(title="Histogram of time to action (checklist_SPCA)")

actions %>% ggplot(mapping=aes(x=checklist_TR,col=animal_type,fill=animal_type)) +
  geom_histogram(binwidth=2,col="red",fill="green",alpha=0.2) +
  coord_cartesian(xlim=c(0,25)) +
  facet_grid(animal_type ~ .) +
  labs(title="Histogram of time to action (checklist_TR)")

actions %>% ggplot(mapping=aes(x=checklist_VET,col=animal_type,fill=animal_type)) +
  geom_histogram(binwidth=2,col="red",fill="green",alpha=0.2) +
  coord_cartesian(xlim=c(0,25)) +
  facet_grid(animal_type ~ .) +
  labs(title="Histogram of time to action (checklist_VET)")

```

